---
title: "examination of weights"
output: html_document
date: "2023-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

## R Markdown

We can view a loess smoother as WLS, in which the weights are proportional to both the distance and the inverse square of the standard error.


```{r echo=T}
source("~/dynamichr/code/utils.R")
source("~/multistate2//code/smoothtest.R")
source("~/multistate2//code/newsmooth.R")
source("../code/fitarray.R")
load("~/Library/CloudStorage/Dropbox-Personal///pheno_dir/output/merged_pheno_censor_final_withdrugs_smoke.rds")
dfh$cad.prs.lec=cut(dfh$cad.prs,breaks = c(-5,-0.84,0.84,5),labels = c("low","mid","high"))
dfh$int=interaction(dfh$f.31.0.0,dfh$cad.prs.lec)
# Relabel the levels of the interaction variable
levels(dfh$int) <- c(1,2,3,4,5,6)


train=dfh[1:(nrow(dfh)*0.80),]
ages=c(40:80)
nstates=c("Health", "Ht","HyperLip","Dm","Cad","death","Ht&HyperLip","HyperLip&Dm","Ht&Dm","Ht&HyperLip&Dm")

#f=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+statin_now+antihtn_now+yearsinstate")
f=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke")


start="Health"
stop="Cad"

c=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = f,window_width = 30,span = 0.75,degree = 2)

ggplotly(c$plot)


```

Here we use the ggplot version which smoothes after expanding x by a grid

```{r}
c2=coefplotsmooth(ages = ages,start = start,stop = stop,modelfit = f)

ggplotly(c2$plot)
```

Here we look at the transition from Ht to Cad and note that some of the noisier coefficients are smoothed;

```{r echo=F}
ages=c(40:80)


start="Ht"
stop="Cad"
c=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = modelfit,window_width = 20,span = 0.5,degree = 2)

ggplotly(c$plot)

```

You can see that the places with the largest errors contribute the least:

```{r echo=F}
head(c$errors[,"statin_now"])
```


How do we use?

1) Create a matrix of coefficients extracted using smoothing for a given model fit and smoothing parameters. Here we'll use a smaller modelfit.
```{r,echo=T}
modelfit=readRDS("~/multistate2/output/fixedsmoke.rds")
start="Health"
stop="Cad"
##
mat=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = modelfit,window_width = 10,span = 0.5,degree = 2)$custom_smooth

head(mat)
```

2) create test data

```{r}
source("../code/fitarray.R")
dfascvd=readRDS("../output/dfascvd_newbp.rds")

fhtrest=merge(dfh,dfascvd[,-which(names(dfascvd)%in%c("age","anylipidmed0","bp_med2","smoke"))],by.x="identifier",by.y="sample_id")


test=dfh[!(dfh$identifier%in%train$identifier),]
test=merge(test,dfascvd[,-which(names(dfascvd)%in%c("age","anylipidmed0","bp_med2","smoke"))],by.x="identifier",by.y="sample_id")
test$ascvd_10y_accaha=test$as2
test$phenos.enrollment=test$f.21003.0.0
test=data.table(test)
```

3) compute prediction product for individuals of those combinations for a given year interval. This will return per year survival and risk probabilities as well as $1 - \prod(per yearsurvival).$

```{r}

#statin==0&anti_htn==0&
modelfit=fitfunc(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke")

atriskmat=data.table(atriskmat)
ar=data.frame(data.table(atriskmat)[smoke==0&anti_htn==0&statin==0,])
mat=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = f,window_width = 20,span = 0.75,degree = 2)$custom_smooth
age=50
c=compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = mat)
attributes(c)

```


Compare coefficients with old approach:

```{r,echo=T,eval=T}
a=return_smoothedmatrix(start = "Ht",stop = "Cad",ages = 40:80,modelfit = abinom)
ggplotly(coefplotsmooth(start = "Ht",stop = "Cad",ages = 40:80,modelfit = abinom)$plot)
b=coefplotsmooth2(ages = ages,start = "Ht",stop = "Cad",modelfit = abinom,window_width = 15,span = 0.75,degree = 2)
ggplotly(b$plot)

```

```{r}
par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$coefmat[,i+1],b$mat[,i],xlab="GGPlotLOESSnoWeights",ylab="WeightedSmoothing",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$original[,i],a$coefmat[,(i+1)],xlab="Original",ylab="LoessSmoothed",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$original[,i],b$mat[,i],xlab="Original",ylab="WLS",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

```



I think it must be the smoothing within the ggplot that works so robustly. They create a number of Xs for each actual X...


Now let's do it a bit differently and compute for actual people:

```{r}

modelfit=fitfunc(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke")

####
ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(levels(test$int)))
lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(levels(test$int)))

b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)

coefs=b$custom_smooth

head(coefs)

for(i in 1:length(agesint)){
  age=agesint[i]
for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]

    atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age &
                        age < HyperLip_0_censor_age &
                        age < Dm_0_censor_age&int==cat, ]
  ar=data.frame("intercept"=1,atrisk$cad.prs,atrisk$f.31.0.0,atrisk$smoke)
  ten.year.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = coefs)$PredictedIntervalrisk)
  lifetime.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(80)),coefmat = coefs)$PredictedIntervalrisk)
  
}}


columns_to_check <- c("Cad_0_censor_age", "Ht_0_censor_age", "HyperLip_0_censor_age", "Dm_0_censor_age")
operators_to_use <- c("<", "<", "<", "<")
int_value_to_filter <- "some_value"  # Replace with appropriate value

empresult = empriskforloop(test = test, columns = columns_to_check, operators = operators_to_use, agesint = agesint, prs_quants = prs_quants)

ascvd.ten.year=ascvdriskforloop(test = test,columns = columns_to_check,operators = operators_to_use,agesint = agesint,prs_quants = prs_quants)

pcten=gofRMSE(Obs = as.vector(empresult$emp.ten.year*100),Prd = as.vector(ascvd.ten.year))
msten=gofRMSE(Obs = as.vector(empresult$emp.ten.year*100),Prd = as.vector(ten.year.new*100))
plotfuncrmse(ascvd.ten.year = ascvd.ten.year,emp.ten.year = emp.ten.year,mstate.ten.year = ten.year.new)

> msten
[1] 0.547
> pcten
[1] 5.518
```

Now try with a fuller model:

```{r}

modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+antihtn_now") 

#pcten 0.679

# modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0")
#pcten 0.592



#modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke") 0.611
#modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="f.31.0.0") 2.162 pcten

####
#ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
#lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)
coefs=b$custom_smooth
agesint=seq(40,70,by=5)
prs_quants=c(data.frame(test%>%group_by(int)%>%summarize(median(cad.prs),mean(cad.prs)))[c(2,4,6),3])

ascvd.ten.year=ascvd.30.year=emp.ten.year=emp.lifetime=ten.year.new=lifetime.new=pce.reverse.tenyear=ascvd.30.year.rc=data.frame()

for(i in 1:length(agesint)) {
  age = agesint[i]
  for (j in 1:length(levels(test$int))) {
    cat = levels(test$int)[j]
    
    atrisk = test[age < Cad_0_censor_age &
                    age < Ht_0_censor_age &
                    age < HyperLip_0_censor_age &
                    age < Dm_0_censor_age & int == cat, ]
    ar = data.frame(
      "intercept" = 1,
      "cad.prs" = atrisk$cad.prs,
      "sex" = atrisk$f.31.0.0,
      "smoke" = atrisk$smoke,
      "antihtn_now" = ifelse(atrisk$antihtn == 1 &
                               atrisk$htn_age < age, 1, 0)
    )
    
    # ar = data.frame(
    #   "intercept" = 1,
    #   "cad.prs" = atrisk$cad.prs,
    #   "sex" = atrisk$f.31.0.0,
    #   "smoke" = atrisk$smoke
    # )
    #   ar = data.frame(
    #   "intercept" = 1,
    #   #"cad.prs" = atrisk$cad.prs,
    #   "sex" = atrisk$f.31.0.0
    #   #"smoke" = atrisk$smoke
    # )
    #ar=data.frame("intercept"=1,"cad.prs"=atrisk$cad.prs,"sex"=atrisk$f.31.0.0)
    ten.year.new[i, j] = mean(
      compute_prediction_product_matrix(
        atrisk = ar,
        agepredinterval = c(age:(age + 10)),
        coefmat = coefs
      )$PredictedIntervalrisk
    )
    lifetime.new[i, j] = mean(
      compute_prediction_product_matrix(
        atrisk = ar,
        agepredinterval = c(age:(80)),
        coefmat = coefs
      )$PredictedIntervalrisk
    )
    emp.ten.year[i, j] = compute_empiricalrisk(age = age,
                                               age2 = age + 10,
                                               atrisk = atrisk)
    emp.lifetime[i, j] = compute_empiricalrisk(age = age,
                                               age2 = 80,
                                               atrisk = atrisk)
    
    ascvd.ten.year[i, j] = compute_pce_predictedrisk(age = age, atrisk = atrisk)## wtih people who are that age at enrollement
    
    dat = data.frame(
      "id" = atrisk$identifier,
      "mysex" = as.factor(atrisk$sex),
      "myage" = rep(age, nrow(atrisk)),
      "mysbp" = atrisk$sbp,
      "mytreat" = ifelse(atrisk$antihtn == 1 &
                           atrisk$htn_age < age, 1, 0),
      "mysmoking" = atrisk$smoke,
      "mydiabetes" = ifelse(atrisk$Dm_0_Any == 2 &
                              atrisk$Dm_0_censor_age < age, 1, 0),
      "mytotalchol" = atrisk$choladj,
      "myhdl" = atrisk$hdladj,
      "Race" = atrisk$Race,
      "mystatnow" = ifelse(atrisk$statin == 1 & atrisk$statin_age < age, 1, 0)
    )
    #
    frs = fun.frs_30y(
      dat,
      id = "id",
      sex = "mysex",
      age = "myage",
      sbp = "mysbp",
      treat = "mytreat",
      smoking = "mysmoking",
      diabetes = "mydiabetes",
      totalchol = "mytotalchol",
      hdl = "myhdl"
    )
    ascvd.30.year[i, j] = median(frs$frs_orig)
    ascvd.30.year.rc[i, j] = median(frs$frs_recali)
    pce.reverse.tenyear[i, j] = mean(
      compute_CVrisk2(
        df = dat,
        scores = "as2",
        age = "myage",
        gender = "mysex",
        race = "Race",
        totchol = "mytotalchol",
        sbp = "mysbp",
        hdl = "myhdl",
        bp_med = "mytreat",
        diabetes = "mydiabetes",
        smoker = "mysmoking",
        lipid_med = "mystatnow"
      )$as2
    )
  }
}

library(ehaGoF)
emp.ten.year=as.matrix(emp.ten.year)
emp.lifetime=as.matrix(emp.lifetime)
ten.year.new=as.matrix(ten.year.new)
lifetime.new=as.matrix(lifetime.new)
ascvd.ten.year=as.matrix(ascvd.ten.year)
ascvd.30.year=as.matrix(ascvd.30.year)
ascvd.30.year.rc=as.matrix(ascvd.30.year.rc)
pce.reverse.tenyear=as.matrix(pce.reverse.tenyear)

pcten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ascvd.ten.year))
msten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ten.year.new*100))
pcten2=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(pce.reverse.tenyear))

pclife=gofRMSE(Obs = as.vector(emp.lifetime[c(1:2),]*100),Prd = as.vector(ascvd.30.year[c(1:2),]))
mslife=gofRMSE(Obs = as.vector(emp.lifetime[c(1:2),]*100),Prd = as.vector(lifetime.new[c(1:2),]*100))
pclife2=gofRMSE(Obs = as.vector(emp.lifetime[c(1:2),]*100),Prd = as.vector(ascvd.30.year.rc[c(1:2),]))


a1=plotfuncrmse(ascvd.ten.year = ascvd.30.year,emp.ten.year = emp.lifetime,mstate.ten.year = lifetime.new)$m+ggtitle("RMSE 30 year")

a2=plotfuncrmse(ascvd.ten.year = ascvd.30.year.rc,emp.ten.year = emp.lifetime,mstate.ten.year = lifetime.new)$m+ggtitle("RMSE 30 year,recalibrated")

library(ggpubr)
saveRDS(ggarrange(a1,a2,common.legend = TRUE,legend = "right"),file = "sharedrmse.rds")

#sex only 5.341 and 2.162

##with antihtn
# 0.679 for ten year, 0.593 for life


#wihout antihtn
#0.611 for ten year, 2.23 for life

#without smoke and antithn
#0.592 for ten, 2.1 for life


df=data.frame("sex.only"=2.1,"sex.genetics"=0.592,"sex.genetics.smoking"=0.611,"sex.genetics.smoking.antihtn"=0.679,"pce.all"=pcten2,"pce.enroll"=pcten)

r=melt(df)

g1=ggplot(r,aes(variable,value,fill=variable))+geom_bar(stat="identity")+labs(x="Model",y="RMSE, 10 year score",fill="Score")+theme_classic()+theme(axis.text.x = element_text(face = "bold",size = 12, angle = 90))

saveRDS(g1,file = "../output/rmsemodelten.rds")

dflife=data.frame("sex.only"=5.341,"sex.genetics"=2.1,"sex.genetics.smoking"=2.23,"sex.genetics.smoking.antihtn"=0.593,"frs.all"=pclife,"frs.recal"=pclife2)

r=melt(dflife)

g2=ggplot(r,aes(variable,value,fill=variable))+geom_bar(stat="identity")+labs(x="Model",y="RMSE, Lifetime score",fill="Score")+theme_classic()+theme(axis.text.x = element_text(face = "bold",size = 12, angle = 90))

saveRDS(g2,file = "../output/rmsemodellife.rds")
```

NOw ROC

```{r}

set.seed(456)
enrollments = c(41:70)
aucmat = matrix(NA, nrow = length(enrollments), ncol = 4)
prcmat = matrix(NA, nrow = length(enrollments), ncol = 4)
semat = matrix(NA, nrow = length(enrollments), ncol = 4)

ages = 40:80
enrollments = c(41:70)
## return a matrix of coefficients over all ages for a given state to state transition
b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)
coefs=b$custom_smooth



for (z in 16:length(enrollments)) {
  age = enrollments[z]
  start = age
  stop = 80
  
  df_frame = test
  atrisk = df_frame[age < Cad_0_censor_age &
                      age < Ht_0_censor_age &
                      age < HyperLip_0_censor_age &
                      age < Dm_0_censor_age & smoke == 0 , ]
  
  
  df_updated = data.frame(atrisk)
  ar = data.frame(
    "intercept" = 1,
    "cad.prs" = atrisk$cad.prs,
    "sex" = atrisk$f.31.0.0,
    "smoke" = atrisk$smoke,
    "antihtn_now" = ifelse(atrisk$antihtn == 1 &
                             atrisk$htn_age < age, 1, 0)
  )
  dat = data.frame(
    "id" = atrisk$identifier,
    "mysex" = as.factor(atrisk$sex),
    "myage" = rep(age, nrow(atrisk)),
    "mysbp" = atrisk$sbp,
    "mytreat" = ifelse(atrisk$antihtn == 1 &
                         atrisk$htn_age < age, 1, 0),
    "mysmoking" = atrisk$smoke,
    "mydiabetes" = ifelse(atrisk$Dm_0_Any == 2 &
                            atrisk$Dm_0_censor_age < age, 1, 0),
    "mytotalchol" = atrisk$choladj,
    "myhdl" = atrisk$hdladj,
    "Race" = atrisk$Race,
    "mystatnow" = ifelse(atrisk$statin == 1 &
                           atrisk$statin_age < age, 1, 0)
  )
  
  mso = compute_prediction_product_matrix(
    coefmat = coefs,
    atrisk = ar,
    agepredinterval = c(start:stop)
  )
  
  df_updated$ms = mso$PredictedIntervalrisk
  f = fun.frs_30y(
    dat,
    id = "id",
    sex = "mysex",
    age = "myage",
    sbp = "mysbp",
    treat = "mytreat",
    smoking = "mysmoking",
    diabetes = "mydiabetes",
    totalchol = "mytotalchol",
    hdl = "myhdl"
  )
  df_updated$frs.30y = f$frs_orig
  df_updated$frs.recal = f$frs_recali
  ### return matrix of smoothed coefficeints
  #library(purrr)
  
  rm(atrisk)
  
  #require(pROC)
  df_updated$outcome = ifelse(df_updated$Cad_0_Any == 2 &
                                df_updated$Cad_0_censor_age < stop,
                              1,
                              0)
  #d = df_updated[round(phenos.enrollment, 0) == age, ]
  d = df_updated
  
  #d=d[!is.na(d$ascvd_10y_accaha),]
  aucmat[z, 1] = roc(d$outcome ~ d$ms)$auc
  semat[z, 1] = sqrt(var(roc(d$outcome ~ d$ms)))
  aucmat[z, 2] = roc(d$outcome ~ d$frs.30y)$auc
  semat[z, 2] = sqrt(var(roc(d$outcome ~ d$frs.30y)))
  aucmat[z, 3] = roc(d$outcome ~ d$frs.recal)$auc
  semat[z, 3] = sqrt(var(roc(d$outcome ~ d$frs.recal)))
  aucmat[z, 4] = roc(d$outcome ~ d$cad.prs)$auc
  semat[z, 4] = sqrt(var(roc(d$outcome ~ d$cad.prs)))
  
  
  require(PRROC)
  fg <- d$ms[d$outcome == 1]
  bg <- d$ms[d$outcome == 0]
  
  prcmat[z, 1] = pr.curve(scores.class0 = fg, scores.class1 = bg)$auc.integral
  
  #semat[i,1]=roc(d$outcome~d$ms)$se
  
  #require(PRROC)
  fg <- na.omit(d$frs.30y[d$outcome == 1])
  bg <- na.omit(d$frs.30y[d$outcome == 0])
  prcmat[z, 2] = pr.curve(scores.class0 = fg, scores.class1 = bg)$auc.integral
  
  fg <- na.omit(d$cad.prs[d$outcome == 1])
  bg <- na.omit(d$cad.prs[d$outcome == 0])
  prcmat[z, 3] = pr.curve(scores.class0 = fg, scores.class1 = bg)$auc.integral
  
  fg <- na.omit(d$frs.recal[d$outcome == 1])
  bg <- na.omit(d$frs.recal[d$outcome == 0])
  prcmat[z, 4] = pr.curve(scores.class0 = fg, scores.class1 = bg)$auc.integral
  
  
  print(paste0("Completedforage", age))
  
  
  
}

saveRDS(aucmat,file = "../output/auc.rds")
saveRDS(prcmat,file = "../output/prc.rds")
improv = mean(aucmat[, 1] - aucmat[, 2]) * 100
mprov = mean(aucmat[, 1] - aucmat[, 3]) * 100
rownames(aucmat) = enrollments
m = melt(aucmat, id.vars = "Age")
names(m) = c("Age", "Model", "AUC")
m$Model = as.factor(m$Model)

levels(m$Model)[1] = "MSGene"
levels(m$Model)[2] = "FRS30y"
levels(m$Model)[3] = "FRS30recal"
levels(m$Model)[4] = "PRS only"
m$se=as.vector(semat)
#m = m[m$Model %in% c("MSGene", "FRS30y", "PRS only"), ]
aucplot <-
  ggplot(m[m$Age<55,], 
    aes(x = Age, y = AUC, color = Model, ymin = AUC + se, ymax = AUC - se))+ geom_point() +
  geom_line(aes(group = Model, color = Model), linewidth = 3) +
  geom_pointrange() +
  ylim(0.5, 0.8) +
 theme_classic() + ggtitle(paste0("30-year risk prediction, ROC AUC"))

library(pROC)
rownames(prcmat) = enrollments
m = melt(prcmat, id.vars = "Age")
names(m) = c("Age", "Model", "AUC")
m$Model = as.factor(m$Model)

levels(m$Model)[1] = "MSGene"
levels(m$Model)[2] = "FRS30y"
levels(m$Model)[3] = "FRS30.recal"
levels(m$Model)[3] = "PRSonly"
m$se=as.vector(semat[,c(1,2,3,4)])/2
#m = m[m$Model %in% c("MSGene", "FRS30y", "PRS only"), ]
prplot <-
  ggplot(m[m$Age<55,], 
    aes(x = Age, y = AUC, color = Model, ymin = AUC + se, ymax = AUC - se))+ geom_point() +
  geom_line(aes(group = Model, color = Model), linewidth = 3) +
  geom_pointrange() +
 
 theme_classic() + ggtitle(paste0("30-year risk prediction, Precision Recall"))


ggarrange(aucplot,prplot,common.legend = T,legend = "right")
saveRDS(ggarrange(aucplot,prplot,common.legend = T,legend = "right"),"../output/jointaucprc.rds")
```


```{R}
#Hyperlip
####
ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
b=coefplotsmooth2(ages = ages,start = "HyperLip",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)
coefs=b$custom_smooth

for(i in 1:length(agesint)){
  age=agesint[i]
for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]

    atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age & #Ht_0_Any==2 
                        age > HyperLip_0_censor_age & HyperLip_0_Any==2 &
                        age < Dm_0_censor_age&int==cat, ]
  ar=data.frame("intercept"=1,"cad.prs"=atrisk$cad.prs,"sex"=atrisk$f.31.0.0,"smoke"=atrisk$smoke,
                "antihtn_now"=ifelse(atrisk$antihtn==1&atrisk$htn_age<age,1,0))
  ten.year.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = coefs)$PredictedIntervalrisk)
  lifetime.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(80)),coefmat = coefs)$PredictedIntervalrisk)
  emp.ten.year[i,j]=compute_empiricalrisk(age = age,age2 = age+10,atrisk = atrisk)
  ascvd.ten.year[i,j]=compute_pce_predictedrisk(age = age,atrisk = atrisk)
}}

pcten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ascvd.ten.year))
msten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ten.year.new*100))

```

DM

```{r}
modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+antihtn_now") 

####
ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
b=coefplotsmooth2(ages = ages,start = "Dm",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)
coefs=b$custom_smooth

for(i in 1:length(agesint)){
  age=agesint[i]
for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]

    atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age & #Ht_0_Any==2 
                        age > Dm_0_censor_age & Dm_0_Any==2 &
                        age < HyperLip_0_censor_age&int==cat, ]
  ar=data.frame("intercept"=1,"cad.prs"=atrisk$cad.prs,"sex"=atrisk$f.31.0.0,"smoke"=atrisk$smoke,
                "antihtn_now"=ifelse(atrisk$antihtn==1&atrisk$htn_age<age,1,0))
  ten.year.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = coefs)$PredictedIntervalrisk)
  lifetime.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(80)),coefmat = coefs)$PredictedIntervalrisk)
  emp.ten.year[i,j]=compute_empiricalrisk(age = age,age2 = age+10,atrisk = atrisk)
  ascvd.ten.year[i,j]=compute_pce_predictedrisk(age = age,atrisk = atrisk)
}}

pcten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ascvd.ten.year))
msten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ten.year.new*100))

```

# plot again

```{r}

modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+antihtn_now") 

intercept=1
cad.prs=c(-3,-2,-1,0,1,2)
sex=c(0,1)
smoke=c(0,1)
antihtn_now=c(0,1)

atrisk=expand.grid(intercept,cad.prs,sex,smoke,antihtn_now)
atrisk=atrisk[atrisk$Var4==0&atrisk$Var5==0,]
#atrisk=expand.grid(intercept,cad.prs,sex,smoke)
#atrisk=atrisk[atrisk$Var4==0,]
tenlifeplotting(start = "Health",stop = "Cad",modelfit = modelfit,agesmooth = c(40:80),agesint = c(40:70),atrisk = data.frame(atrisk),window_width = 20,span = 0.75,degree = 2,prs_quants = cad.prs)

start="Health"
stop="Cad"
c=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
b=compute_prediction_product_matrix(atrisk = atrisk,agepredinterval = 40:80,coefmat = c$custom_smooth)
a=apply(data.frame(b$Survival),1,function(x){(1-cumprod(x))})

a2=a
colnames(a2)=c(paste0(rep("F:",6),c(-3:2)),paste0(rep("M:",6),c(-3:2)))
rownames(a2)=ages
m=melt(a2)
ggplot(m,aes(Var1,value,col=Var2))+geom_smooth()+theme_classic()+labs(y="Hazard Starting from 40",x="Ages")

a=apply(data.frame(b$Survival),1,function(x){(1-cumprod(x))})
a2=a
colnames(a2)=c(paste0(rep("F:",6),c(-3:2)),paste0(rep("M:",6),c(-3:2)))
rownames(a2)=ages
m=melt(a2)
ggplot(m,aes(Var1,value,col=Var2))+geom_smooth()+theme_classic()+labs(y="Survival Starting from 40",x="Ages")


