---
title: "examination of weights"
output: html_document
date: "2023-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

## R Markdown

We can view a loess smoother as WLS, in which the weights are proportional to both the distance and the inverse square of the standard error.


```{r echo=T}
source("~/dynamichr/code/utils.R")
source("~/multistate2//code/smoothtest.R")
source("~/multistate2//code/newsmooth.R")
source("../code/fitarray.R")
load("../../Dropbox/pheno_dir/output/merged_pheno_censor_final_withdrugs_smoke.rds")
dfh$cad.prs.lec=cut(dfh$cad.prs,breaks = c(-5,-0.84,0.84,5),labels = c("low","mid","high"))
dfh$int=interaction(dfh$f.31.0.0,dfh$cad.prs.lec)
# Relabel the levels of the interaction variable
levels(dfh$int) <- c(1,2,3,4,5,6)


train=dfh[1:(nrow(dfh)*0.80),]
ages=c(40:80)
nstates=c("Health", "Ht","HyperLip","Dm","Cad","death","Ht&HyperLip","HyperLip&Dm","Ht&Dm","Ht&HyperLip&Dm")

f=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+statin_now+antihtn_now+yearsinstate")


start="Health"
stop="Cad"

c=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = f,window_width = 30,span = 0.75,degree = 2)

ggplotly(c$plot)


```

Here we use the ggplot version which smoothes after expanding x by a grid

```{r}
c2=coefplotsmooth(ages = ages,start = start,stop = stop,modelfit = f)

ggplotly(c2$plot)
```

Here we look at the transition from Ht to Cad and note that some of the noisier coefficients are smoothed;

```{r echo=F}
ages=c(40:80)


start="Ht"
stop="Cad"
c=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = modelfit,window_width = 20,span = 0.5,degree = 2)

ggplotly(c$plot)

```

You can see that the places with the largest errors contribute the least:

```{r echo=F}
head(c$errors[,"statin_now"])
```


How do we use?

1) Create a matrix of coefficients extracted using smoothing for a given model fit and smoothing parameters. Here we'll use a smaller modelfit.
```{r,echo=T}
modelfit=readRDS("~/multistate2/output/fixedsmoke.rds")
start="Health"
stop="Cad"
##
mat=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = modelfit,window_width = 10,span = 0.5,degree = 2)$custom_smooth

head(mat)
```

2) create a matrix with at risk parameters of interest:

```{r}
source("../code/fitarray.R")
test=dfh[!(dfh$identifier%in%train$identifier),]

prs_quants=c(data.frame(test%>%group_by(int)%>%summarize(median(cad.prs),mean(cad.prs)))[c(2,4,6),3])
statin=c(0,1)
antihtn=c(0,1)
sex=c(0,1)
smoke=c(0,1)
# Use expand.grid to create all possible combinations

```

3) compute prediction product for individuals of those combinations for a given year interval. This will return per year survival and risk probabilities as well as $1 - \prod(per yearsurvival).$

```{r}

#statin==0&anti_htn==0&
modelfit=fitfunc(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke")

atriskmat=data.table(atriskmat)
ar=data.frame(data.table(atriskmat)[smoke==0&anti_htn==0&statin==0,])
mat=coefplotsmooth2(ages = ages,start = start,stop = stop,modelfit = f,window_width = 20,span = 0.75,degree = 2)$custom_smooth
age=50
c=compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = mat)
attributes(c)

```

Now we plot:

```{r plotting,echo=F,eval=T}

```



Compare coefficients with old approach:

```{r,echo=T,eval=T}
abinom=readRDS("~/multistate2/output/allcovariates.rds")
a=return_smoothedmatrix(start = "Ht",stop = "Cad",ages = 40:80,modelfit = abinom)
ggplotly(coefplotsmooth(start = "Ht",stop = "Cad",ages = 40:80,modelfit = abinom)$plot)

b=coefplotsmooth2(ages = ages,start = "Ht",stop = "Cad",modelfit = abinom,window_width = 15,span = 0.75,degree = 2)
ggplotly(b$plot)

```

```{r}
par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$coefmat[,i+1],b$mat[,i],xlab="GGPlotLOESSnoWeights",ylab="WeightedSmoothing",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$original[,i],a$coefmat[,(i+1)],xlab="Original",ylab="LoessSmoothed",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

par(mfrow=c(3,3))
for(i in c(1:7)){plot(a$original[,i],b$mat[,i],xlab="Original",ylab="WLS",main=paste0("Variable_",names(a$coefmat)[i+1]));abline(c(0,1),col="red")}

```



I think it must be the smoothing within the ggplot that works so robustly. They create a number of Xs for each actual X...


Now let's do it a bit differently and compute for actual people:

```{r}

modelfit=fitfunc(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke")

####
ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)

b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)

coefs=b$custom_smooth

head(coefs)

for(i in 1:length(agesint)){
  age=agesint[i]
for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]

    atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age &
                        age < HyperLip_0_censor_age &
                        age < Dm_0_censor_age&int==cat, ]
  ar=data.frame("intercept"=1,atrisk$cad.prs,atrisk$f.31.0.0,atrisk$smoke)
  ten.year.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = coefs)$PredictedIntervalrisk)
  lifetime.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(80)),coefmat = coefs)$PredictedIntervalrisk)
}}


emp.ten.year=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
emp.lifetime=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
test=data.table(test)

for(i in 1:length((agesint))){
  age=agesint[i]
  for(j in 1:length(levels(test$int))){
    
    cat=levels(test$int)[j]
  atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age &
                        age < HyperLip_0_censor_age &
                        age < Dm_0_censor_age&int==cat
               
                , ] 
    print(dim(atrisk))
    emp.ten.year[i,j]=compute_empiricalrisk(age=age,age2 = age+10,atrisk = atrisk)
    emp.lifetime[i,j]=compute_empiricalrisk(age=age,age2 = 100,atrisk  = atrisk)
  }}


test=dfh[!(dfh$identifier%in%train$identifier),]
ascvd.ten.year=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
dfascvd=readRDS("../output/dfascvd_newbp.rds")
test=merge(test,dfascvd[,c("sample_id","as2")],by.x="identifier",by.y="sample_id")
test$ascvd_10y_accaha=test$as2
test$phenos.enrollment=test$f.21003.0.0
test=data.table(test)
for(i in 1:length((agesint))){
  age=agesint[i]
  for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]
  atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age &
                        age < HyperLip_0_censor_age &
                        age < Dm_0_censor_age&int==cat 
                #&smoke==0
                #&(antihtn==0|htn_age>age)&(statin==0|statin_age>age)
                , ] 
    ascvd.ten.year[i,j]=compute_pce_predictedrisk(age=age,atrisk = atrisk)
    #ascvdriskmat[i,2]=compute_empiricalrisk(age=40,df_frame = mpce,cat = cat)
  }
}


pcten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ascvd.ten.year))
msten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ten.year.new*100))

plotfuncrmse(ascvd.ten.year = ascvd.ten.year,emp.ten.year = emp.ten.year,mstate.ten.year = ten.year.new)


```

NOw try with a fuller model:

```{r}

modelfit=fitfunc(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+antihtn_now+statin_now")

####
ten.year.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)
lifetime.new=matrix(NA,nrow = length(agesint),ncol=length(prs_quants)*2)

b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
ggplotly(b$plot)

coefs=b$custom_smooth



for(i in 1:length(agesint)){
  age=agesint[i]
for(j in 1:length(levels(test$int))){
    cat=levels(test$int)[j]

    atrisk = test[age < Cad_0_censor_age &
                        age < Ht_0_censor_age &
                        age < HyperLip_0_censor_age &
                        age < Dm_0_censor_age&int==cat, ]
  ar=data.frame("intercept"=1,"cad.prs"=atrisk$cad.prs,"sex"=atrisk$f.31.0.0,"smoke"=atrisk$smoke,
                "antihtn_now"=ifelse(atrisk$antihtn==1&atrisk$htn_age<age,1,0),"statin_now"=ifelse(atrisk$statin==1&atrisk$statin_age<age,1,0))
  ten.year.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(age+10)),coefmat = coefs)$PredictedIntervalrisk)
  lifetime.new[i,j]=mean(compute_prediction_product_matrix(atrisk = ar,agepredinterval = c(age:(80)),coefmat = coefs)$PredictedIntervalrisk)
}}


plotfuncrmse(ascvd.ten.year = ascvd.ten.year,emp.ten.year = emp.ten.year,mstate.ten.year = ten.year.new)


msten=gofRMSE(Obs = as.vector(emp.ten.year*100),Prd = as.vector(ten.year.new*100))
# 0.866
```

