---
title: "flashier_nmg"
output: html_document
date: "2023-12-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```


## The EBMF Model

Empirical Bayes matrix factorization (Wang and Stephens, 2021) decomposes a data matrix $Y \in \mathbb{R}^{n \times p}$ as
$$ Y = LF' + E, $$
with "loadings" $L \in \mathbb{R}^{n \times K}$, "factors" $F \in \mathbb{R}^{p \times K}$, and residual errors
$$ e_{ij} \sim N(0, 1 / \tau_{ij}).$$
The model puts priors on each factor and each set of loadings:
$$f_{\cdot k} \sim g^{(f)}_k,\ \ell_{\cdot k} \sim g^{(\ell)}_k,$$
with $g^{(\ell)}_k$ and $g^{(f)}_k$ assumed to belong to some families of distributions $\mathcal{G}^{(\ell)}$ and $\mathcal{G}^{(f)}$ and estimated using the data. The default choice of prior family for both factors and loadings is the family of point-normal distributions:
$$ g \sim \pi_0 \delta_0 + (1 - \pi_0) N(0, \sigma^2), $$
where both $\pi_0$ and $\sigma^2$ are free parameters. This family is especially useful when factors and loadings can be expected to be at least modestly sparse; in other settings, different prior families might be preferred.

To avoid over-parametrization, it is necessary to make some assumptions about the precision parameters $\tau_{ij}$. The default assumption is that all $\tau_{ij}$s are equal:
$$ e_{ij} \sim N(0, 1 / \tau).$$

Note that when the prior families $\mathcal{G}^{(\ell)}$ and $\mathcal{G}^{(f)}$ are closed under scaling (as is typically the case), then the model as formulated above is not identifiable, since we can scale each set of loadings $\ell_{\cdot k}$ by any constant $d_k \ne 0$ if we also scale factor $f_{\cdot k}$ by $1 / d_k$. 

We can make the matrix factorization unique by writing
$$ Y = LDF' + E, $$
with the scales of loadings $\ell_{\cdot 1}, \ldots, \ell_{\cdot K}$ and factors $f_{\cdot 1}, \ldots, f_{\cdot K}$ constrained in some fashion (for example, by requiring $\| \ell_{\cdot k} \|_2 = 1$ and $\| f_{\cdot k} \|_2 = 1$ for all $k$). We refer to such a factorization as an **LDF factorization**.

# Age up to and including

Here we will examine the sparse factor analysis of individuals who receive a code at the age under each plot indicated

```{r,eval=F,echo=TRUE}
remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
library(flashier)
library(data.table)
library(Matrix)
filepath="~/Library/CloudStorage/Dropbox-Personal/pheno_dir/"
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 5)
## for age 40 we specify that 
for(i in 1:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
print(age_threshold)
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",age_threshold,".rds"))
}
```

Now let's plot:

```{r,echo=F}
age_cutoffs <- seq(40, 90, by = 5)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = TRUE), n = 10),
            las = 2,
            main = paste("Age leq", age_threshold))
  }
}



```


# All ages

We can also do for all, but we'll need to first remove rows with 0s and binarize the matrix such as non-zero values are 1 and zero values are 0.

Recall we shall keep a non-negative prior on the factors and loadings,

```{r,eval=F,echo=T}
library(Matrix)
library(flashier)
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)
#install.packages("flashier")
X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
#K=10
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS_930.rds")
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS.rds")


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_norestrictions.rds")
rm(f2)


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=0)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_s0.rds")
rm(f2)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=1)
saveRDS(f2,"~/Library/CloudStorage/Dropbox-Personal/factor_run47_s1.rds")

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,S=0)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run10_s0.rds")
rm(f2)

```

Here we make no restriction and examine the top 20 variable loadings:

```{r echo=T}
## this one was with the colSums must be >500 and no specification on S
f = readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_run10.rds")
par(mfrow=c(3,3))
for(k in 2:10){
barplot(head(sort(f[["F_pm"]][,k],decreasing = TRUE),n=10),las=2)
}

```


# Group by age
## looking at the codes that occur exactly at that age group

```{r,eval=F,echo=T}

#remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 10)
## for age 40 we specify that 
for(i in 1:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
min=ifelse(i==1,0,age_cutoffs[i-1])
print(c(min,age_threshold))
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & min < x & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,s=1/1500)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",age_threshold,".rds"))
}
```

Now let's plot:

```{r echo=F}
age_cutoffs <- seq(40, 90, by = 10)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = TRUE), n = 10),
            las = 2,
            main = paste0("Age", age_threshold))
  }
}

```

# Wth fixed s0 and k=10 
*  why does mean come out when K = 10 but not K=47
* the mean has been removed?
```{r echo=T}
#f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,S=0)
#saveRDS(f2,file= "~/Library/CloudStorage/Dropbox-Personal/factor_run47.rds")
f=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run47_s0.rds")
plot(f)
head(sort(f$F_pm[,1],decreasing = TRUE),n=10)
```



```{r,echo=FALSE,eval=FALSE}
## Show the pchiatric

l=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run47_s0.rds")

plot(l)

barplot(sort(l$F_pm[,25],decreasing = TRUE)[1:20],las=2)

apply(df.phecode.wide[,..w],2,function(x){median(na.omit(x))})
```

Let's try with a very small s and show that ELBO worsens.


```{r echo=T}
# df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode//df.phecode.wide.rds")
# dfw=df.phecode.wide[,-c(1:5)]
# dfw[!is.na(dfw)]=1
# dfw[is.na(dfw)]=0
# X=as.matrix(dfw)
# 
# X=X[,which(colSums(X)>0)]
# X=X[which(rowSums(X)>0),]
# X=Matrix(X,sparse = T)

# smin=min(f$residuals_sd)
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=smin)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run48_smin.rds")
f2=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run48_smin.rds")
f2$elbo
```

However, the elbo is worse here than not specifying s

```{r}
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=1/sqrt(ncol(X)))
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run50_sqrtncol.rds")
f2=readRDS("~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run50_sqrtncol.rds")
f2$elbo


#f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
#saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run50_nospec.rds")
f2=readRDS("~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run50_nospec.rds")
f2$elbo
```



Show that perhaps the point exponential forces cardiac away and weird removal of factors; here we usee point exponential. So now we will use normal mis (no restrictions)

```{r}
filepath="~/Library/CloudStorage/Dropbox-Personal/pheno_dir/"

## with no positivity constraint
fit=readRDS(paste0(filepath,"fit_default_all.rds"))

fit$F_ghat[[1]]

head(sort(fit$F_pm[,2],decreasing = TRUE,n=20))
tail(sort(fit$F_pm[,2],decreasing = TRUE,n=20))

head(sort(fit$F_pm[,3],decreasing = TRUE,n=20))
tail(sort(fit$F_pm[,3],decreasing = TRUE,n=20))

head(sort(fit$F_pm[,4],decreasing = TRUE,n=20))
tail(sort(fit$F_pm[,4],decreasing = TRUE,n=20))

## with positivity constraint
# flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
# Nullchecking 50 factors...
# Factor1set to zero, increasing objective by 1.045e+07.
# Factor5set to zero, increasing objective by 9.593e+05.
# Factor11set to zero, increasing objective by 5.946e+05.

f=readRDS("~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run50_nospec.rds")
lapply(seq(1:20),function(x){head(sort(f$F_pm[,x],decreasing = TRUE),n=10)})

## problem here is that the cardiovascular effect is the mean effect (1) and i thought we should remove this?
## but elbo is maximal
f$elbo
```

## Here we do on the arrays


We initialize a (tensor) of patients times traits x epochs (5 year intervals from 40-80). The advantage here is we can use rows with 0s as the code allows. There are some features in the tensors we might prefer, but constraining to be all positive has some odd results - esophageal diseases appear in several factors with high loadings, as does pregnancy.

```{r}

# array
# dfw[is.na(dfw)]=0
# dat=as.matrix(dfw)
# rm(dfw)
# age_cutoffs <- seq(40, 90, by = 5)
# tdat <- lapply(age_cutoffs, function(cutoff) Matrix(ifelse(dat > 0 & dat < cutoff,1,0), sparse = TRUE))
# class(tdat) <- "sptensor"

#f=flashier::flash(data = tdat,ebnm_fn = c(ebnm_point_exponential,ebnm_point_laplace,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
#saveRDS(f,file = "~/Library/CloudStorage/Dropbox-Personal/factor_array_10_exponential.rds")
# f2=flashier::flash(data = tdat,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_array_20_allpos.rds")

f2=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/factor_array_20_allpos.rds")
lapply(seq(1:10),function(x){head(sort(f2$loadings_pm[[2]][,x],decreasing = TRUE),n=20)})

```

We use  ebnm_point_exponential,ebnm_point_laplace,ebnm_point_exponential on total array

Hwwever, now heart disease is 4 and 5 
```{r}
f=readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_array_10_exponential.rds")
lapply(seq(1:10),function(x){head(sort(f$loadings_pm[[2]][,x],decreasing = TRUE),n=20)})
```

With  all positive array (all point_exponential)

```{r}
f=readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_array_20_allpos.rds")
lapply(seq(1:10),function(x){head(sort(f$loadings_pm[[2]][,x],decreasing = TRUE),n=10)})
```


### Let's return to the matrix space

We can redo  removing only 0 row, not remoivng null rows or look if we set S to something small.
```{r echo=T,eval=F}
remotes::install_github("willwerscheid/flashier", ref = "allow_zero_rows",force=T)
library(flashier)
library(Matrix)
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode//df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)
X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20,nullcheck = F,backfit = T)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit.rds")


f3=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,nullcheck = T,backfit = T)
#saveRDS(f3,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_nc.rds")


```

First we try with a normal loadings constraint so that we don't have this strange behavior of removal of the first factor, but it seems that then the first factor is the only that contains Htn and ischemic heart disease.

```{}
#f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20,nullcheck = F,backfit = T)
f2=readRDS("~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit.rds")
lapply(seq(1:20),function(x){head(sort(f2$F_pm[,x],decreasing = TRUE),n=10)})
```


Let's do with null check =T and we see that heart disease appears again inthe 2nd,3rd and 4th with high loadings 
```{r}
#f3=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,nullcheck = T,backfit = T)
f3=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_nc.rds")

lapply(seq(1:20),function(x){head(sort(f3$F_pm[,x],decreasing = TRUE),n=10)})
```

It seems we got better results with no restrictions (default normal) so let's try with backfitting.

```{r}
#f3=flashier::flash(data = X,var_type = 2,greedy_Kmax = 50,nullcheck = T,backfit = T)
#saveRDS(f3,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz.rds")
f3=readRDS("~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz.rds")
lapply(seq(1:40),function(x){head(sort(f3$F_pm[,x],decreasing = TRUE),n=10)})
```

But now how do we interpret negative factors, and ischemic hear disease in 4,7,11...


Ok what if we make the loadings flexible but non negaitve and allow for 0 rows (with new branch update) and backfit

```{r}
#f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,nullcheck = F,backfit = T,S=1/1500)
#saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/matrix_pn_pe_backfit_nonneg_allrows.rds")

f2=readRDS("~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz_allrows.rds")
lapply(seq(1:20),function(x){head(sort(f2$F_pm[,x],decreasing = TRUE),n=10)})
```
So is the ischemic factor 2 or 20? And why does GERD enter everywhere ...