---
title: "flashier_nmg"
output: html_document
date: "2023-12-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Age up to and including

Here we will examine the sparse factor analysis of individuals who receive a code at the age under each plot indicated

```{r,eval=F}
#remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
library(flashier)
library(data.table)
library(Matrix)
filepath="~/Library/CloudStorage/Dropbox-Personal/pheno_dir/"
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 5)
## for age 40 we specify that 
for(i in 1:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
print(age_threshold)
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",age_threshold,".rds"))
}
```

Now let's plot:

```{r}
age_cutoffs <- seq(40, 90, by = 5)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = T), n = 20),
            las = 2,
            main = paste("Age leq", age_threshold))
  }
}

```


# All ages

We can also do for all, but we'll need to first remove rows with 0s and binarize the matrix such as non-zero values are 1 and zero values are 0.

Recall we shall keep a non-negative prior on the factors and loadings,

```{r,eval=F}
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)
#install.packages("flashier")
X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
#K=10
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS_930.rds")
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS.rds")


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_norestrictions.rds")
rm(f2)


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=0)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_s0.rds")
rm(f2)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=1)
saveRDS(f2,"~/Library/CloudStorage/Dropbox-Personal/factor_run47_s1.rds")
```

Here we make no restriction and examine the top 20 variable loadings:

```{r}
## this one was with the colSums must be >500 and no specification on S
f = readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_run10.rds")
par(mfrow=c(3,3))
for(k in 2:10){
barplot(head(sort(f[["F_pm"]][,k],decreasing = T),n=20),las=2)
}

```


# Group by age
## looking at the codes that occur exactly at that age group

```{r,eval=F}
#remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 10)
## for age 40 we specify that 
for(i in 2:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
min=ifelse(i==1,0,age_cutoffs[i-1])
print(c(min,age_threshold))
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & min < x & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",age_threshold,".rds"))
}
```

Now let's plot:

```{r}
age_cutoffs <- seq(40, 90, by = 10)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = T), n = 20),
            las = 2,
            main = paste0("Age", age_threshold))
  }
}

```

# Wth fixed s0
```{r}
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,S=0)
file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47.rds")
head(sort(f2$F_pm[,2],decreasing = T))