---
title: "flashier_nmg"
output: html_document
date: "2023-12-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## The EBMF Model

Empirical Bayes matrix factorization (Wang and Stephens, 2021) decomposes a data matrix $Y \in \mathbb{R}^{n \times p}$ as
$$ Y = LF' + E, $$
with "loadings" $L \in \mathbb{R}^{n \times K}$, "factors" $F \in \mathbb{R}^{p \times K}$, and residual errors
$$ e_{ij} \sim N(0, 1 / \tau_{ij}).$$
The model puts priors on each factor and each set of loadings:
$$f_{\cdot k} \sim g^{(f)}_k,\ \ell_{\cdot k} \sim g^{(\ell)}_k,$$
with $g^{(\ell)}_k$ and $g^{(f)}_k$ assumed to belong to some families of distributions $\mathcal{G}^{(\ell)}$ and $\mathcal{G}^{(f)}$ and estimated using the data. The default choice of prior family for both factors and loadings is the family of point-normal distributions:
$$ g \sim \pi_0 \delta_0 + (1 - \pi_0) N(0, \sigma^2), $$
where both $\pi_0$ and $\sigma^2$ are free parameters. This family is especially useful when factors and loadings can be expected to be at least modestly sparse; in other settings, different prior families might be preferred.

To avoid over-parametrization, it is necessary to make some assumptions about the precision parameters $\tau_{ij}$. The default assumption is that all $\tau_{ij}$s are equal:
$$ e_{ij} \sim N(0, 1 / \tau).$$

Note that when the prior families $\mathcal{G}^{(\ell)}$ and $\mathcal{G}^{(f)}$ are closed under scaling (as is typically the case), then the model as formulated above is not identifiable, since we can scale each set of loadings $\ell_{\cdot k}$ by any constant $d_k \ne 0$ if we also scale factor $f_{\cdot k}$ by $1 / d_k$. 

We can make the matrix factorization unique by writing
$$ Y = LDF' + E, $$
with the scales of loadings $\ell_{\cdot 1}, \ldots, \ell_{\cdot K}$ and factors $f_{\cdot 1}, \ldots, f_{\cdot K}$ constrained in some fashion (for example, by requiring $\| \ell_{\cdot k} \|_2 = 1$ and $\| f_{\cdot k} \|_2 = 1$ for all $k$). We refer to such a factorization as an **LDF factorization**.

# Age up to and including

Here we will examine the sparse factor analysis of individuals who receive a code at the age under each plot indicated

```{r,eval=F}
remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
library(flashier)
library(data.table)
library(Matrix)
filepath="~/Library/CloudStorage/Dropbox-Personal/pheno_dir/"
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 5)
## for age 40 we specify that 
for(i in 1:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
print(age_threshold)
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",age_threshold,".rds"))
}
```

Now let's plot:

```{r}
age_cutoffs <- seq(40, 90, by = 5)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = T), n = 20),
            las = 2,
            main = paste("Age leq", age_threshold))
  }
}

```


# All ages

We can also do for all, but we'll need to first remove rows with 0s and binarize the matrix such as non-zero values are 1 and zero values are 0.

Recall we shall keep a non-negative prior on the factors and loadings,

```{r,eval=F}
library(Matrix)
library(flashier)
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)
#install.packages("flashier")
X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
#K=10
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS_930.rds")
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_withoutfixingS.rds")


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_norestrictions.rds")
rm(f2)


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=0)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run47_s0.rds")
rm(f2)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=1)
saveRDS(f2,"~/Library/CloudStorage/Dropbox-Personal/factor_run47_s1.rds")

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,S=0)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run10_s0.rds")
rm(f2)

```

Here we make no restriction and examine the top 20 variable loadings:

```{r}
## this one was with the colSums must be >500 and no specification on S
f = readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_run10.rds")
par(mfrow=c(3,3))
for(k in 2:10){
barplot(head(sort(f[["F_pm"]][,k],decreasing = T),n=20),las=2)
}

```


# Group by age
## looking at the codes that occur exactly at that age group

```{r,eval=F}
#remotes::install_github("willwerscheid/flashier", ref = "sparse_tensors")
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode/df.phecode.wide.rds")
#df.phecode.wide[is.na(df.phecode.wide)]=0
dfw=df.phecode.wide[,-c(1:5)]
age_cutoffs <- seq(40, 90, by = 10)
## for age 40 we specify that 
for(i in 1:length(age_cutoffs)){
dat=df.phecode.wide[,-c(1:5)]
age_threshold=age_cutoffs[i]
min=ifelse(i==1,0,age_cutoffs[i-1])
print(c(min,age_threshold))
dat[, (1:ncol(dat)) := lapply(.SD, function(x) ifelse(!is.na(x) & min < x & x < age_threshold, 1, 0))]
#install.packages("flashier")

X=as.matrix(dat)
#install.packages("flashier")
## remove individuals never phenotyped
## for 40 and 45: 
X=X[,which(colSums(X)>0)]
#X=X[,which(colSums(X)>100)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
print(dim(X))
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,s=1/1500)
rm(X)
rm(dat)
saveRDS(f2,paste0("~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",age_threshold,".rds"))
}
```

Now let's plot:

```{r}
age_cutoffs <- seq(40, 90, by = 10)
K = 10
for (i in 1:length(age_cutoffs)) {
  age_threshold = age_cutoffs[i]
  f = readRDS(
    paste0(
      "~/Library/CloudStorage/Dropbox-Personal/new_flash_by_age/flash_age_exact",
      age_threshold,
      ".rds"
    )
  )
  par(mfrow = c(3, 3))
  for (k in 2:f$n_factors) {
    barplot(head(sort(f[["F_pm"]][, k], decreasing = T), n = 20),
            las = 2,
            main = paste0("Age", age_threshold))
  }
}

```

# Wth fixed s0 and k=10 
*  why does mean come out when K = 10 but not K=47
```{r}
#f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10,S=0)
#saveRDS(f2,file= "~/Library/CloudStorage/Dropbox-Personal/factor_run47.rds")
f=readRDS(file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run47_s0.rds")
plot(f2)
f2=head(sort(f2$F_pm[,2],decreasing = T))

```

For Jordan:

```{r}
plot(l)

barplot(sort(l$F_pm[,25],decreasing = TRUE)[1:20],las=2)

apply(df.phecode.wide[,..w],2,function(x){median(na.omit(x))})
```

Let's try with a very small s
```{r}

df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode//df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)

X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)
f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)

# smin=min(f$residuals_sd)
# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=smin)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run48_smin.rds")
# ## but the elbo is worse here than not specifying s


# f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,S=1/sqrt(ncol(X)))
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run50_sqrtncol.rds")


f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
#saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_run50_nospec.rds")

# array
# dfw[is.na(dfw)]=0
# dat=as.matrix(dfw)
# rm(dfw)
# age_cutoffs <- seq(40, 90, by = 5)
# tdat <- lapply(age_cutoffs, function(cutoff) Matrix(ifelse(dat > 0 & dat < cutoff,1,0), sparse = TRUE))
# class(tdat) <- "sptensor"

#f=flashier::flash(data = tdat,ebnm_fn = c(ebnm_point_exponential,ebnm_point_laplace,ebnm_point_exponential),var_type = 2,greedy_Kmax = 10)
#saveRDS(f,file = "~/Library/CloudStorage/Dropbox-Personal/factor_array_10_exponential.rds")
# f2=flashier::flash(data = tdat,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20)
# saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/factor_array_20_allpos.rds")
```

Show that perhaps the point exponential forces cardiac away and weird removal of factors:

```{r}
filepath="~/Library/CloudStorage/Dropbox-Personal/pheno_dir/"

## with no positivity constraint
fit=readRDS(paste0(filepath,"fit_default_all.rds"))

fit$F_ghat[[1]]

head(sort(fit$F_pm[,2],decreasing = T,n=20))
tail(sort(fit$F_pm[,2],decreasing = T,n=20))

head(sort(fit$F_pm[,3],decreasing = T,n=20))
tail(sort(fit$F_pm[,3],decreasing = T,n=20))

head(sort(fit$F_pm[,4],decreasing = T,n=20))
tail(sort(fit$F_pm[,4],decreasing = T,n=20))

## with positivity constraint
# flashier::flash(data = X,ebnm_fn = c(ebnm_point_exponential,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50)
# Nullchecking 50 factors...
# Factor1set to zero, increasing objective by 1.045e+07.
# Factor5set to zero, increasing objective by 9.593e+05.
# Factor11set to zero, increasing objective by 5.946e+05.

#f=readRDS("~/Library/CloudStorage/Dropbox-Personal/flash_allages/factor_run50_nospec.rds")
lapply(seq(1:10),function(x){head(sort(f$F_pm[,x],decreasing = T),n=20)})



## with ebnm_point_exponential,ebnm_point_laplace,ebnm_point_exponential
f=readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_array_10_exponential.rds")
lapply(seq(1:10),function(x){head(sort(f$loadings_pm[[2]][,x],decreasing = T),n=20)})

## with  all positive array

f=readRDS("~/Library/CloudStorage/Dropbox-Personal/factor_array_20_allpos.rds")
lapply(seq(1:10),function(x){head(sort(f$loadings_pm[[2]][,x],decreasing = T),n=10)})

#### redo without removing
remotes::install_github("willwerscheid/flashier", ref = "allow_zero_rows",force=T)
library(flashier)
library(Matrix)
df.phecode.wide=readRDS("~/Library/CloudStorage/Dropbox-Personal/phecode//df.phecode.wide.rds")
dfw=df.phecode.wide[,-c(1:5)]
dfw[!is.na(dfw)]=1
dfw[is.na(dfw)]=0
X=as.matrix(dfw)
X=X[,which(colSums(X)>0)]
X=X[which(rowSums(X)>0),]
X=Matrix(X,sparse = T)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20,nullcheck = F,backfit = T)
saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit.rds")


f3=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,nullcheck = T,backfit = T)
#saveRDS(f3,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_nc.rds")

lapply(seq(1:10),function(x){head(sort(f3$F_pm[,x],decreasing = T),n=20)})


f3=flashier::flash(data = X,var_type = 2,greedy_Kmax = 50,nullcheck = T,backfit = T)
saveRDS(f3,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz.rds")
f3=readRDS("~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz.rds")
lapply(seq(1:40),function(x){head(sort(f3$F_pm[,x],decreasing = TRUE),n=10)})

### make the loadings flexible but non negaitve
## backfit
### clone branch with s very samll so i don't have to (allow 0 rows)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 20,nullcheck = F,backfit = T,S=1/1500)

saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz_allrows.rds")



### make all non negaitve
## backfit
### clone branch with s very samll so i don't have to (allow 0 rows)

f2=flashier::flash(data = X,ebnm_fn = c(ebnm_point_normal,ebnm_point_exponential),var_type = 2,greedy_Kmax = 50,nullcheck = F,backfit = T,S=1/1500)

saveRDS(f2,file = "~/Library/CloudStorage/Dropbox-Personal/flash_allages/matrix_pn_pe_backfit_nonneg_allrows.rds")

f2=readRDS("~/Library/CloudStorage/Dropbox-Personal/matrix_pn_pe_backfit_norestrictionz_allrows.rds")
lapply(seq(1:20),function(x){head(sort(f2$F_pm[,x],decreasing = TRUE),n=10)})
```
