---
title: "RCA Analysis"
output:
  html_document: default
  pdf_document: default
date: '2023-08-29'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,eval = T,tidy = T,warning = FALSE,message = FALSE)
```

## Congenital RCA Anomalous Coronaries

Here we analyse the age/sex matched controls with the 187 cases provided.
We remove select cases and produce three compariosns:

* Treated versus untreated RCA cases have no difference in mortality
* When comparing among unroofing, CABG, cath or no treatment, no differences in mortality
* when comparing to healthy age and sex matched controls no difference

```{r cars}
library('data.table')
library(lubridate)
library(MatchIt)
library(survival)
library(survminer)
library(dplyr)

ada=fread("~/Library/CloudStorage/Dropbox-Personal/AnomalousCoronaryArt_DATA_2023-08-29_0921.csv")
ada=ada[ada$type_of_anomaly==3,]

ada$censor.age=as.numeric(difftime(ada$date_of_death,ada$dob,units = "days")/365.25)
##match on white

alive=which(ada$alive==1)
## get age of censor for those that live
ada$censor.age[alive]=difftime(ada$last_date_of_follow_up[alive],ada$dob[alive],units = "days")/365.25

## there are a few without censorship times
ada=ada[-which(is.na(ada$censor.age)),]
ada=ada[-which(is.na(ada$race)),]


```


Now let's answer our questions:

#1) survival between surgical unroofing or reimplantation vs. cabg vs. cath vs. no intervention.

First, check if any intervention is useful:

```{r}

ada$death=ifelse(ada$alive==1,0,1)


fit=survfit(Surv(time = censor.age,event = death)~intervention,data=ada)


fit

ggsurvplot(
  fit,
  data = ada,
  size = 1,                 # change line size
  palette =
    c("#E7B800", "#2E9FDF"),# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  risk.table.col = "strata",# Risk table color by groups
  legend.labs =
    c("Not","Intervention"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw()      # Change ggplot2 theme
)
```


### Now let's do for the different types of intervention versus none:



```{r}
ada$treat=NULL
ada$treat[ada$intervention_list==2|ada$intervention_list==3]="U"
ada$treat[ada$intervention_list==1]="B"
ada$treat[ada$the_type_of_intervention==1]="C"
ada$treat[ada$intervention==0]="none"
ada$treat=factor(ada$treat,levels=c("U","B","C","none"),labels=c("unroofing_implant","bypass","cath","none"))
ada=ada[-which(is.na(ada$treat)),]
fit=survfit(Surv(time = censor.age,event = death)~treat,data=ada)


fit

ggsurvplot(
  fit,
  data = ada,
  size = 1,                 # change line size
    conf.int = TRUE,          # Add confidence interval
  pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  risk.table.col = "strata",# Risk table color by groups
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw()      # Change ggplot2 theme
)

```

However, when we break by intervention, we see that there is a difference among the 4 categories of patients:

```{r}
ada$race=ifelse(ada$race==3,1,0)
ada$sex=ifelse(ada$sex==2,1,0)
## all.equal(which(is.na(ada$censor.age)),which(ada$alive==1))
ada$dm=ifelse(ada$comorbidities___1==1,1,0)
ada$smoke=ifelse(ada$comorbidities___2==1,1,0)
ada$hyperlip=ifelse(ada$comorbidities___3==1,1,0)
ada$hypertension=ifelse(ada$comorbidities___4==1,1,0)
ada$stroke=ifelse(ada$comorbidities___5==1,1,0)

ada_subset=ada[,c("record_id","sex","dob","alive","censor.age","hyperlip","hypertension","stroke","dm","smoke","race","treat")]

ada_subset%>%group_by(treat)%>%summarise("Per_female"=mean(sex),"hyperlip"=mean(hyperlip),"stroke"=mean(stroke),"dm"=mean(dm),"smoke"=mean(smoke),"race"=mean(race),"birthyear"=mean(year(dob)))
```


### Now:

And then a comparison of the low surgical risk (who do not have slit-like, and do not have intramural course) vs healthy controls?

```{r}

ada_subset=ada[,c("record_id","sex","dob","alive","censor.age","hyperlip","hypertension","stroke","dm","smoke","race","intervention")]

f=fread("~/Library/CloudStorage/Dropbox-PartnersHealthCare/MGBB (1)/Phenotypes/Manual Curated/MGBB_Phenos_2CODE 2022-06-21.csv")

f$hyperlip=ifelse(f$Has_hyperlipidemia==1,1,0)
f$stroke=ifelse(f$Has_all_stroke_noTIA==1,1,0)
f$hypertension=ifelse(f$Has_htn==1,1,0)

cov=fread("~/Library/CloudStorage/Dropbox-PartnersHealthCare/cathproject/mgbb_pce_collection_date.csv")
f=merge(f,cov,by="EMPI")
f$smoke=ifelse(f$current_smoking_imputed=="yes",1,0)
f$dm=ifelse(f$dm=="yes",1,0)
f$race=ifelse(f$race_imputed=="white",1,0)

## remove those without a date of death
f2=f[!which(f$Vital_Status=="Deceased"&is.na(f$Date_of_Death)),]

f2$censor.age=as.numeric(difftime(f2$Date_of_Death,f2$Date_of_Birth,units = "days")/365.25)

# w=which(f2$Vital_Status=="Living")
# n=which(is.na(f2$censor.age))
# all.equal(w,n)
alive=which(f2$Vital_Status=="Living")
f2$censor.age[alive]=difftime(f2$Censor_Date[alive],f2$Date_of_Birth[alive],units = "days")/365.25

#########

f2$treat=rep(2,nrow(f2))
f2$Sex=ifelse(f2$sex_imputed=="female",1,0)
f2$Vital_Status=ifelse(f2$Vital_Status=="Living",1,0)

f2_subset=f2[,c("Biobank_Subject_ID","Sex","Date_of_Birth","Vital_Status","censor.age","hyperlip","hypertension","stroke","dm","smoke","race","treat")]



names(f2_subset)=names(ada_subset)

df=rbind(ada_subset,f2_subset)
df$age.today=as.numeric(difftime(as.Date("2023-08-26"),as.Date(df$dob),units = "days")/365.25)
df$death=ifelse(df$alive==0,1,0)
df$case_control=ifelse(df$intervention==2,0,1)
m.out0 <- matchit(case_control ~ age.today + sex, data = df,
                 method = NULL, distance = "glm")

summary(m.out0)
```



##  Now do nearest neighbor propensity score matching:

Here we display info about the matching after controlling for DOB, sex matching all those in the RCA group (both treated and untreated) with health controls:

```{r pressure, echo=FALSE}
m.out1 <- matchit(case_control ~ age.today+sex, data = df,method = "nearest", distance = "glm")

m.out1
summary(m.out1, un = FALSE)

#check the balance
plot(m.out1, type = "density", interactive = FALSE,
     which.xs = ~age.today+sex)

```


Now let's generate the survival model:

```{r}
library(survival)
library(survminer)

m=match.data(m.out1)
m$intervention=factor(m$intervention,levels = c(0,1,2),labels=c("untreated","treated","control"))
fit=survfit(Surv(time = censor.age,event = death)~intervention,data=m)
fit

g=ggsurvplot(
  fit,
  data = m,
  size = 1,                 # change line size
# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  risk.table.col = "strata",# Risk table color by groups
 
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw()      # Change ggplot2 theme
)

g


```


