---
title: "MoreSelective"
output: html_document
date: "2023-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

## let's be more selective and select first and last script date

* Here we grab the first script date and the last script date for treated individuals

```{r}
source("~/multistate2/code/utils.R")
library(MatchIt)
prs_subset=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/prs_subset.rds")
#glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s.txt")
df_baseline=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/dfukb_baseline.rds")

glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s_withi.txt")
colnames(glp1)[1]=c("Identifier")
colnames(glp1)[3]=c("event_dt")
glp1$event_dt <- as.Date(glp1$event_dt, format="%d/%m/%Y")


biomarkers=readRDS("~/Library/CloudStorage/Dropbox-Personal/ukbb-ehr-data/data/biomarkers.rds")
#bmi=biomarkers[biomarkers$variable%in%"bmi",]

bmi=biomarkers[biomarkers$variable%in%"weight",]
bmi$date <- as.Date(bmi$date, format="%Y/%m/%d")


patient_med_data <- glp1 %>%
  group_by(Identifier) %>%
  summarize(first_script_date = min(event_dt),last_script_date = max(event_dt))

colnames(patient_med_data)=c("eid","first_script_date","last_script_date")
patient_med_data$first_script_date <- as.Date(patient_med_data$first_script_date, format = "%Y-%m-%d")
patient_med_data$last_script_date <- as.Date(patient_med_data$last_script_date, format = "%Y-%m-%d")
```

## Treated Folks

* We grab the first BMI as the median BMI that is within 1 year of the first script date
* We grab the post-treat BMI as the median BMI that is within 12 weeks after 1st script to 1 year after last.
* We tried the last measurement before but it is noisy to use the final measurement only, we can also see how many measurements folks had

```{r}
patient_bmi_data=bmi[bmi$eid%in%patient_med_data$eid,c("eid","date","value")]
bbmi=df_baseline[,c("identifier","f.53.0.0","f.21001.0.0")]
bbmi$f.53.0.0=as.Date(bbmi$f.53.0.0)

patient_bmi_data=rbind(patient_bmi_data,bbmi[,c("identifier","f.53.0.0","f.21001.0.0")],use.names=FALSE)

patient_bmi_data$date <- as.Date(patient_bmi_data$date, format = "%Y-%m-%d")
merged_data <- merge(patient_bmi_data, patient_med_data, by = "eid")

bmi_before_script <- merged_data %>%
  filter(date < first_script_date & date > (first_script_date - years(1)))%>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"medbmi"=median(value),"number"=length(value),)

summary(bmi_before_script$number)

bmi_after_script <- merged_data %>%
  filter(date > (first_script_date + weeks(12)) & date < (last_script_date + years(1))) %>%
  
  group_by(eid) %>%
  summarise("afterdate"=min(date),"medbmi"=median(value),"number"=length(value))

summary(bmi_after_script$number)
```

## Now we merge the first and last BMI measurements with the script data and perform checks

* to make sure first measure is before drug start
*  check to make sure last BMI measure is at least 12 weeks after drug start

```{r,echo=T}
final_data <- (bmi_before_script[,c("eid","beforedate","medbmi")]) %>%
  left_join(patient_med_data[,c("eid","first_script_date","last_script_date")], by = "eid") %>%
  left_join(bmi_after_script[,c("eid","afterdate","medbmi")], by = "eid")
colnames(final_data)=c("eid","pretreat_measure_date","pretreat_bmi","first_script_date","last_script_date","post_drug_measure_date","post_drug_bmi")

final_data=na.omit(final_data)

summary(year(final_data$pretreat_measure_date))
summary(year(final_data$post_drug_measure_date))

# check to make sure first measure is before drug start
sum(difftime(final_data$first_script_date,final_data$pretreat_measure_date)<0)

sum(difftime(final_data$pretreat_measure_date,final_data$post_drug_measure_date,units = "days")==0)

# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data$post_drug_measure_date,final_data$first_script_date,units = "days")<83)
## controls

```

# we will choose controls to mirror our selection above

## This is less clearcut for controls so ...
* First BMI median between the treated populations pretreat measure dates (min,max pre_treat_date)
* Last BMI median between the treated populations post-treat measure dates

```{r}
controls=bmi[!bmi$eid%in%patient_med_data$eid,]

## since there is no med, grab first and last measruement 
## find measruements around the time of treated

bmi_controls_first <- controls %>%
  filter(date >= min((final_data$pretreat_measure_date)) & date < max((final_data$pretreat_measure_date))) %>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"meanbmi"=median(value),"number"=length(value))


summary(bmi_controls_first$beforedate)
summary(bmi_controls_first$number)

bmi_controls_last <- controls %>%
  filter(date >= min(final_data$post_drug_measure_date) & date < max(final_data$post_drug_measure_date))  %>%
  group_by(eid) %>%
 summarise("afterdate"=min(date),"meanbmi"=median(value),"number"=length(value))

summary(bmi_controls_last$afterdate)
summary(bmi_controls_last$number)

final_data_controls=
  inner_join(bmi_controls_first[,c("eid","beforedate","meanbmi")], bmi_controls_last[,c("eid","afterdate","meanbmi")],by = "eid", suffix = c("_before", "_after"))

# remote those in which dates are the same

final_data_controls=final_data_controls[final_data_controls$beforedate!=final_data_controls$afterdate,]


```

Again, checks:


```{r,echo=T}
# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")>0)
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")==0)


## treated
summary(year(final_data$pretreat_measure_date))
summary(year(final_data$post_drug_measure_date))

## controls
summary(year(final_data_controls$beforedate))
summary(year(final_data_controls$afterdate))

final_data$treat=1
final_data_controls$treat=0

treat=final_data[,c("eid","pretreat_bmi","post_drug_bmi","pretreat_measure_date","post_drug_measure_date","treat","first_script_date")]
untreat=final_data_controls[,c("eid","meanbmi_before","meanbmi_after","beforedate","afterdate","treat","beforedate")]
names(untreat)=names(treat)
all_data=rbind(treat,untreat)
all_data$time_diff=difftime(all_data$post_drug_measure_date,all_data$pretreat_measure_date,units = "days")

df=merge(all_data,df_baseline[,c("identifier","f.34.0.0","f.31.0.0")],by.x="eid",by.y="identifier")
df_with_prs=merge(df,prs_subset[,c("Identifier","BMI")],by.x="eid",by.y="Identifier")

```

## now we perform matching on

* duration between measurements
* patient birthday
* sex
* pretreat BMI

```{r}
## now match on pretreat bmi
df_with_prs$bmi.prs=df_with_prs$BMI
matcher=matchit(treat~pretreat_bmi+f.34.0.0+f.31.0.0+time_diff+bmi.prs,data = df_with_prs)
summary(matcher)

#extract matched data
match_dat=match.data(matcher)

match_dat=rename(match_dat,c(sex=f.31.0.0,birth_year=f.34.0.0))%>%select(-c(BMI,weights,subclass))
pop.quants=quantile(prs_subset$BMI,probs = seq(0,1,by=0.25))
match_dat$prs.group=cut(match_dat$bmi.prs,breaks = pop.quants,labels = c(1:4))


#eliminate those with CAD at baseline measure
#match_dat=match_dat[match_dat$Cad_censor_age>match_dat$first_bmi_age,]


match_dat$delta=match_dat$post_drug_bmi-match_dat$pretreat_bmi
match_dat$pretreat_weight=match_dat$pretreat_bmi


model=lm(delta~treat+treat:bmi.prs+pretreat_weight+sex+birth_year,match_dat)

#model=lm(delta~treat+treat:bmi.prs,match_dat)


summary(model)
confint.lm(model)
t.test(match_dat$delta[match_dat$treat==1])

t.test(match_dat$delta[match_dat$treat==0])
```



## suppose we want to plot the combined effect of treatment and treatment:BMI PRS

```{r}
prs.quants=quantile(prs_subset$BMI,probs = seq(0,1,by=0.10))
prs.quants=c(-3:3)
coefficients <- coef(model)
cov_matrix <- vcov(model)


# Prepare data frame for the forest plot
# Prepare data frame for the forest plot
forest_data <- data.frame(Quantile = factor(prs.quants, levels = prs.quants), Estimate = NA, SE = NA)

# Loop over the quantile values
for (i in 1:length(prs.quants)) {
    # Calculate combined effect
    combined_effect <- coefficients["treat"] + coefficients["treat:bmi.prs"] * prs.quants[i]

    # Calculate combined variance
    var_treat_bmi_prs_adjusted <- prs.quants[i]^2 * cov_matrix["treat:bmi.prs", "treat:bmi.prs"]
    combined_variance <- cov_matrix["treat", "treat"] + var_treat_bmi_prs_adjusted + 2 * prs.quants[i] * cov_matrix["treat", "treat:bmi.prs"]
    
    # Store in data frame
    forest_data$Estimate[i] <- combined_effect
    forest_data$SE[i] <- sqrt(combined_variance)
}

# Create the forest plot with effect estimate on X-axis and PRS quantile on Y-axis
ggplot(forest_data, aes(x = Estimate, y = Quantile, xmin = Estimate - 1.96 * SE, xmax = Estimate + 1.96 * SE)) +
    geom_pointrange() +
    ylab("BMI.PRS SD") +
    xlab("Combined Effect Estimate") +
    theme_minimal()


```


## Let's do some 3d plots

```{r}
library(plotly)

# Assuming 'data' has 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
fig <- plot_ly(match_dat[match_dat$treat==1,], x = ~pretreat_bmi, y = ~delta, z = ~bmi.prs,
color = ~bmi.prs,
type = "scatter3d", mode = "markers")
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Baseline BMI'),
yaxis = list(title = 'Change in BMI'),
zaxis = list(title = 'PRS')),
title = "3D Plot of Change in BMI by PRS, Baseline BMI, and Treatment")
fig


# Assuming 'data' has 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
fig <- plot_ly(match_dat[match_dat$treat==0,], x = ~pretreat_bmi, y = ~delta, z = ~bmi.prs,
color = ~bmi.prs,
type = "scatter3d", mode = "markers")
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Baseline BMI'),
yaxis = list(title = 'Change in BMI'),
zaxis = list(title = 'PRS')),
title = "3D Plot of Change in BMI by PRS, Baseline BMI, and Treatment")
fig

```

# We examine distribution of delta by treatment and BMI

```{r}
load("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/merged_pheno_censor_final_withdrugs_smoke.rds")
pheno=dfh

match_dat=merge(match_dat,pheno,by.x="eid",by.y="identifier")

match_dat$enrollage=difftime(match_dat$f.53.0.0,match_dat$Birthdate,units = "days")/365.25
match_dat$first_bmi_age=difftime(match_dat$pretreat_measure_date,match_dat$Birthdate,units = "days")/365.25

library(table1)
dat=match_dat
dat$HyperLip_0_Any=factor(dat$HyperLip_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$Ht_0_Any=factor(dat$Ht_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$Dm_0_Any=factor(dat$Dm_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$sex=factor(dat$sex,levels = c(0,1),labels = c("Female","Male"))
dat$enrollage=as.numeric(dat$enrollage)
dat$smoke=factor(dat$smoke,levels=c(0,1),labels = c("No","Yes"))
dat$antihtn=factor(dat$antihtn,levels=c(0,1),labels = c("No","Yes"))
dat$statin=factor(dat$statin,levels=c(0,1),labels = c("No","Yes"))
dat$treat=factor(dat$treat,levels=c(0,1),labels=c("Untreated","Treated"))
label(dat$sex)<- "Sex"
label(dat$enrollage) <- "Age"
label(dat$pretreat_bmi) <- "Pre-Treat BMI"
label(dat$HyperLip_0_Any) = "Hyperlipidemia"
label(dat$Ht_0_Any) = "Hypertension"
label(dat$Dm_0_Any) = "Diabetes Type 2"
dat$treat=as.factor(dat$treat)
label(dat$statin)="Statin"
label(dat$smoke)="Current Smoker"
label(dat$antihtn)="Anti Htn"
# One level of stratification
table1(~sex + enrollage +pretreat_bmi+HyperLip_0_Any+Ht_0_Any+Dm_0_Any+statin+smoke+antihtn|treat,dat)

```



```{r,echo=FALSE,eval=FALSE}

# Now let's plot some stuff

library(ggplot2)
        
ggplot(match_dat,aes(y=delta,x=treat,group=treat,fill=as.factor(treat)))+
  geom_boxplot()+labs(y="Change in Weight (kg) ",x="Treat",fill="Treat")+theme_classic()

ggplot(match_dat,aes(x=as.numeric(time_diff)/7,fill=as.factor(treat)))+geom_density()+
  labs(x="Weeks Between Measurements",main="Density",fill="Treat")+facet_wrap(~treat)+theme_classic()


levels(match_dat$treat) <- c("Untreated", "Treated")
ggplot(match_dat,aes(y=delta,x=prs.group,group=prs.group,fill=as.factor(prs.group)))+
  geom_boxplot()+labs(y="Change in Weight (kg)",x="PRS Quartile (population based)",fill="Prs Quartile")+theme_classic()+facet_wrap(~as.factor(treat),nrow = 1)



# Assuming you have a dataframe 'data' with columns 'baseline_BMI', 'change_in_BMI', and 'PRS'
match_dat$treat=factor(match_dat$treat,levels = c(0,1),labels=c("Untreated","Treated"))
ggplot(match_dat, aes(x = pretreat_bmi, y = delta,col=prs.group)) +
  geom_point() +  # Adjust the geom as necessary
  facet_grid(treat ~ prs.group) +
  labs(title = "Delta vs Baseline BMI by Treatment Status and PRS Quartiles",
       x = "Baseline Wt",
       y = "Delta Wt",
       col="BMI PRS Quartile",
       caption = "PRS Quartiles: 1 (Lowest) to 4 (Highest)") +
  theme_classic()

library(ggplot2)

# Assuming your dataframe is named 'data' with 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
ggplot(match_dat, aes(x = pretreat_bmi, y = delta, size = bmi.prs, color = as.factor(treat))) +
  geom_point(alpha = 0.7) +  # Adjust alpha for bubble transparency if needed
  scale_size_continuous(range = c(1, 10)) +  # Adjust bubble size range as needed
  scale_color_brewer(palette = "Set1") +  # Choose a color palette that suits your data
  labs(title = "Bubble Plot of Change in BMI by Baseline BMI, PRS, and Treatment",
       x = "Baseline Wt ",
       y = "Change in Wt",
       size = "PRS") +
  theme_minimal() +
  guides(color = guide_legend(title = "Treatment Group"), size = guide_legend(title = "PRS"))
```

## And here we summarize by population bmi.prs quartiles:

```{r}
match_dat$treat=factor(match_dat$treat,levels = c(0,1),labels=c("Untreated","Treated"))

match_dat[match_dat$treat=="Untreated",]%>%group_by(prs.group)%>%summarise(median(delta),IQR(delta),sd(delta))


match_dat[match_dat$treat=="Treated",]%>%group_by(prs.group)%>%summarise(median(delta),IQR(delta),sd(delta))

```


# Plot Delta Weight by BMI.prs

```{r}
ggplot(data = match_dat,aes(x=bmi.prs,y=delta,col=as.factor(treat),fill=as.factor(treat)))+geom_point()+stat_smooth()+facet_wrap(~treat)+labs(y="Delta Weight (kg)",x="BMI PRS")+theme_classic()

ggplot(data = match_dat,aes(x=pretreat_bmi,y=delta,col=as.factor(treat),fill=as.factor(treat)))+geom_point()+stat_smooth()+facet_wrap(~treat)+labs(y="Delta Weight (kg)",x="PreTreat BMI")+theme_classic()
```


# Bayesian Prior Analysis
## Now let's use the results from UKB as a prior in an analysis of the GLP1 data in AoU:

# Bayesian Analysis with Conjugate Normals

```{r}
# Prior parameters (from first dataset)
mu_0 <- -0.44 # mean of the prior
se_0 <- 0.196770 # standard error of the prior
sigma_0_squared <- se_0^2 # variance of the prior

# Likelihood parameters (from new dataset)
mu <- -0.26 # mean of the likelihood
se <- 0.15 # standard error of the likelihood
sigma_squared <- se^2 # variance of the likelihood

# Calculating the posterior mean (mu_n) and variance (sigma_n_squared)
mu_n <- (sigma_squared * mu_0 + sigma_0_squared * mu) / (sigma_squared + sigma_0_squared)
sigma_n_squared <- 1 / (1 / sigma_0_squared + 1 / sigma_squared)

# Posterior Standard Deviation
sigma_n <- sqrt(sigma_n_squared)

# Define a sequence of values for x-axis
x_values <- seq(mu_n - 4 * sigma_n, mu_n + 4 * sigma_n, length.out = 1000)

# Compute the density of the normal distribution for posterior
posterior_values <- dnorm(x_values, mean = mu_n, sd = sigma_n)

# Calculate probability of effect less than 0
prob_less_than_zero <- pnorm(0, mean = mu_n, sd = sigma_n)

# Calculate the 95% credible interval
ci_lower <- qnorm(0.025, mean = mu_n, sd = sigma_n)
ci_upper <- qnorm(0.975, mean = mu_n, sd = sigma_n)

# Compute the density of the normal distribution for prior, likelihood, and posterior
prior_values <- dnorm(x_values, mean = mu_0, sd = sqrt(sigma_0_squared))
likelihood_values <- dnorm(x_values, mean = mu, sd = sqrt(sigma_squared))
posterior_values <- dnorm(x_values, mean = mu_n, sd = sqrt(sigma_n_squared))

# Calculate the 95% credible interval for the posterior
ci_lower <- qnorm(0.025, mean = mu_n, sd = sigma_n)
ci_upper <- qnorm(0.975, mean = mu_n, sd = sigma_n)

# Create a data frame for plotting
plot_data <- data.frame(x = x_values, Prior = prior_values, Likelihood = likelihood_values, Posterior = posterior_values)

# Create the plot
ggplot(plot_data, aes(x)) + 
    geom_line(aes(y = Prior, colour = "Prior")) +
    geom_line(aes(y = Likelihood, colour = "Likelihood")) +
    geom_line(aes(y = Posterior, colour = "Posterior")) +
    geom_area(data = subset(plot_data, x < 0), aes(y = Posterior), fill = "red", alpha = 0.3) +
    geom_vline(xintercept = ci_lower, linetype = "dashed", color = "blue") +
    geom_vline(xintercept = ci_upper, linetype = "dashed", color = "blue") +
    scale_colour_manual(values = c("Prior" = "blue", "Likelihood" = "green", "Posterior" = "red")) +
    ggtitle("Comparison of Prior, Likelihood, and Posterior Distributions with Credible Interval") +
    xlab("Interaction Effect Size") +
    ylab("Density") +
    annotate("text", x = mu_n, y = max(posterior_values), label = paste("Posterior Mean:", round(mu_n, 3)), vjust = -1, color = "black") +
    annotate("text", x = ci_lower, y = max(posterior_values)/2, label = paste("95% CI: [", round(ci_lower, 3), ",", round(ci_upper, 3), "]"), vjust = -1, color = "black") +
    theme_classic()
```



## Let's examine other PCS

```{r}
df_with_prs=merge(df_with_prs,prs_subset[,c("Identifier","BC","CAD","T2D","LDL_SF","HT","ISS")],by.x="eid",by.y=
                  "Identifier")
matcher=matchit(treat~pretreat_bmi+f.34.0.0+f.31.0.0+time_diff+BMI,data = df_with_prs)

#extract matched data
match_dat=match.data(matcher)


match_dat$delta=match_dat$post_drug_bmi-match_dat$pretreat_bmi
match_dat$pretreat_weight=match_dat$pretreat_bmi



#eliminate those with CAD at baseline measure
#match_dat=match_dat[match_dat$Cad_censor_age>match_dat$first_bmi_age,]

match_dat=rename(match_dat,c(sex=f.31.0.0,birth_year=f.34.0.0))%>%select(-c(weights,subclass))




prs_cols <- c("BC","CAD","T2D","LDL_SF","HT","ISS") # replace with your actual column names
models <- list() # to store all the models

for (prs_col in prs_cols) {
  # Construct the formula as a string
  formula_str <- paste("delta ~ treat+sex+pretreat_weight+birth_year + treat:", prs_col, sep = "")
  
  # Convert the string to a formula
  formula <- as.formula(formula_str)
  
  # Fit the model using lm()
  model <- lm(formula, data = match_dat)
  
  # Store the model in the list
  models[[prs_col]] <- summary(model)
}

models


```

## summarize by med type

```{r}
glp1_set=glp1[glp1$Identifier%in%match_dat$eid,]
glp1_set$base=str_split_fixed(glp1_set$V7,pattern = " +",n=2)[,1]

df=data.frame(glp1_set%>%group_by(base)%>%summarize(length(base),length(unique(Identifier))))

df=rename(df,"Total Scripts"=length.base.,"Unique Individuals"=length.unique.Identifier..)

m=melt(df)
ggplot(m,aes(x = base,y=value,fill=variable))+
  geom_bar(stat = "identity",position = "dodge")+
  theme_classic()

# Initialize an empty data frame to store results
results <- data.frame(Drug = character(), 
                      PredictedLoss = numeric(), 
                      StdError = numeric())


for(drug in c("Bydureon","Byetta","Dulaglutide","Exenatide","Liraglutide","Lixisenatide","Trulicity","Victoza")){
  print(drug)
  i=glp1_set$Identifier[glp1_set$base%in%drug]
  d=match_dat[match_dat$eid%in%i&match_dat$treat==1,]
  d$sex=as.factor(d$sex)
  fit=lm(delta~bmi.prs+pretreat_weight+sex+birth_year,d)
  print(fit$coefficients)
  print(mean(d$bmi.prs))
  print(mean(d$pretreat_weight))
    mean_values <- data.frame(
    bmi.prs = 1,
    pretreat_weight = mean(d$pretreat_weight, na.rm = TRUE),
    sex = factor(0, levels = levels(d$sex)),  # Assuming '1' is the most common category
    birth_year = mean(d$birth_year, na.rm = TRUE)
  )

 pred <- predict(fit, newdata = mean_values, se.fit = TRUE)
  
  # Add results to the data frame
  results <- rbind(results, data.frame(Drug = drug, 
                                       PredictedLoss = pred$fit, 
                                       StdError = pred$se.fit))
}


ggplot(results, aes(x = Drug, y = PredictedLoss, ymin = PredictedLoss - StdError, ymax = PredictedLoss + StdError)) +
  geom_point() +
  geom_errorbar(width = 0.2) +
  theme_minimal() +
  xlab("Drug Class") +
  ylab("Predicted Weight Loss") +
  ggtitle("Predicted Weight Loss per Drug Class with Standard Errors")
```