---
title: "MoreSelective"
output: html_document
date: "2023-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

## let's be more selective and select first and last script date

* Here we grab the first script date and the last script date for treated individuals

```{r}
source("~/multistate2/code/utils.R")
library(MatchIt)
prs_subset=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/prs_subset.rds")
#glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s.txt")
df_baseline=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/dfukb_baseline.rds")

glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s_withi.txt")
colnames(glp1)[1]=c("Identifier")
colnames(glp1)[3]=c("event_dt")
glp1$event_dt <- as.Date(glp1$event_dt, format="%d/%m/%Y")


biomarkers=readRDS("~/Library/CloudStorage/Dropbox-Personal/ukbb-ehr-data/data/biomarkers.rds")
bmi=biomarkers[biomarkers$variable%in%"bmi",]
#bmi=biomarkers[biomarkers$variable%in%"weight",]
bmi$date <- as.Date(bmi$date, format="%Y/%m/%d")


patient_med_data <- glp1 %>%
  group_by(Identifier) %>%
  summarize(first_script_date = min(event_dt),last_script_date = max(event_dt))

colnames(patient_med_data)=c("eid","first_script_date","last_script_date")
patient_med_data$first_script_date <- as.Date(patient_med_data$first_script_date, format = "%Y-%m-%d")
patient_med_data$last_script_date <- as.Date(patient_med_data$last_script_date, format = "%Y-%m-%d")
```

## Treated Folks

* We grab the first BMI as the median BMI that is within 1 year of the first script date
* We grab the post-treat BMI as the median BMI that is within 12 weeks after 1st script to 1 year after last.
* We tried the last measurement before but it is noisy to use the final measurement only, we can also see how many measurements folks had

```{r}
patient_bmi_data=bmi[bmi$eid%in%patient_med_data$eid,c("eid","date","value")]
bbmi=df_baseline[,c("identifier","f.53.0.0","f.21001.0.0")]
bbmi$f.53.0.0=as.Date(bbmi$f.53.0.0)

patient_bmi_data=rbind(patient_bmi_data,bbmi[,c("identifier","f.53.0.0","f.21001.0.0")],use.names=FALSE)

patient_bmi_data$date <- as.Date(patient_bmi_data$date, format = "%Y-%m-%d")
merged_data <- merge(patient_bmi_data, patient_med_data, by = "eid")

bmi_before_script <- merged_data %>%
  filter(date < first_script_date & date > (first_script_date - years(1)))%>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"medbmi"=median(value),"number"=length(value),)

summary(bmi_before_script$number)

bmi_after_script <- merged_data %>%
  filter(date > (first_script_date + weeks(12)) & date < (last_script_date + years(1))) %>%
  
  group_by(eid) %>%
  summarise("afterdate"=min(date),"medbmi"=median(value),"number"=length(value))

summary(bmi_after_script$number)
```

## Now we merge the first and last BMI measurements with the script data and perform checks
* to make sure first measure is before drug start
*  check to make sure last BMI measure is at least 12 weeks after drug start

```{r}
final_data <- (bmi_before_script[,c("eid","beforedate","medbmi")]) %>%
  left_join(patient_med_data[,c("eid","first_script_date","last_script_date")], by = "eid") %>%
  left_join(bmi_after_script[,c("eid","afterdate","medbmi")], by = "eid")
colnames(final_data)=c("eid","pretreat_measure_date","pretreat_bmi","first_script_date","last_script_date","post_drug_measure_date","post_drug_bmi")

final_data=na.omit(final_data)

summary(year(final_data$pretreat_measure_date))
summary(year(final_data$post_drug_measure_date))

# check to make sure first measure is before drug start
sum(difftime(final_data$first_script_date,final_data$pretreat_measure_date)<0)

# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data$post_drug_measure_date,final_data$first_script_date,units = "days")<83)
## controls

```

# we will choose controls to mirror our selection above

## This is less clearcut for controls so ...
* First BMI median between the treated populations pretreat measure dates (min,max pre_treat_date)
* Last BMI median between the treated populations post-treat measure dates

```{r}
controls=bmi[!bmi$eid%in%patient_med_data$eid,]

## since there is no med, grab first and last measruement 
## find measruements around the time of treated

bmi_controls_first <- controls %>%
  filter(date >= min((final_data$pretreat_measure_date)) & date < max((final_data$pretreat_measure_date))) %>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"meanbmi"=median(value),"number"=length(value))


summary(bmi_controls_first$beforedate)
summary(bmi_controls_first$number)

bmi_controls_last <- controls %>%
  filter(date >= min(final_data$post_drug_measure_date) & date < max(final_data$post_drug_measure_date))  %>%
  group_by(eid) %>%
 summarise("afterdate"=min(date),"meanbmi"=median(value),"number"=length(value))

summary(bmi_controls_last$afterdate)
summary(bmi_controls_last$number)

final_data_controls=
  inner_join(bmi_controls_first[,c("eid","beforedate","meanbmi")], bmi_controls_last[,c("eid","afterdate","meanbmi")],by = "eid", suffix = c("_before", "_after"))

# remote those in which dates are the same

final_data_controls=final_data_controls[final_data_controls$beforedate!=final_data_controls$afterdate,]
```

Again, checks:


```{r}
# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")>0)
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")==0)


## treated
summary(year(final_data$pretreat_measure_date))
summary(year(final_data$post_drug_measure_date))

## controls
summary(year(final_data_controls$beforedate))
summary(year(final_data_controls$afterdate))

final_data$treat=1
final_data_controls$treat=0

treat=final_data[,c("eid","pretreat_bmi","post_drug_bmi","pretreat_measure_date","post_drug_measure_date","treat","first_script_date")]
untreat=final_data_controls[,c("eid","meanbmi_before","meanbmi_after","beforedate","afterdate","treat","beforedate")]
names(untreat)=names(treat)
all_data=rbind(treat,untreat)
all_data$time_diff=difftime(all_data$post_drug_measure_date,all_data$pretreat_measure_date,units = "days")

df=merge(all_data,df_baseline[,c("identifier","f.34.0.0","f.31.0.0")],by.x="eid",by.y="identifier")
df_with_prs=merge(df,prs_subset[,c("Identifier","BMI")],by.x="eid",by.y="Identifier")
```

## now we perform matching on

* duration between measurements
* patient birthday
* sex
* pretreat BMI

```{r}

df_with_prs$bmi.prs=df_with_prs$BMI
matcher=matchit(treat~pretreat_bmi+f.34.0.0+f.31.0.0+time_diff+bmi.prs,data = df_with_prs)
summary(matcher)

#extract matched data
match_dat=match.data(matcher)
match_dat$bmi.prs=match_dat$BMI



match_dat$delta=match_dat$post_drug_bmi-match_dat$pretreat_bmi

model=lm(delta~treat+treat:bmi.prs+pretreat_bmi,match_dat)
summary(model)
```



## suppose we want to plot the combined effect of treatment and treatment:BMI PRS

```{r}
prs.quants=quantile(prs_subset$BMI,probs = seq(0,1,by=0.10))
prs.quants=c(-3:3)
coefficients <- coef(model)
cov_matrix <- vcov(model)


# Prepare data frame for the forest plot
# Prepare data frame for the forest plot
forest_data <- data.frame(Quantile = factor(prs.quants, levels = prs.quants), Estimate = NA, SE = NA)

# Loop over the quantile values
for (i in 1:length(prs.quants)) {
    # Calculate combined effect
    combined_effect <- coefficients["treat"] + coefficients["treat:bmi.prs"] * prs.quants[i]

    # Calculate combined variance
    var_treat_bmi_prs_adjusted <- prs.quants[i]^2 * cov_matrix["treat:bmi.prs", "treat:bmi.prs"]
    combined_variance <- cov_matrix["treat", "treat"] + var_treat_bmi_prs_adjusted + 2 * prs.quants[i] * cov_matrix["treat", "treat:bmi.prs"]
    
    # Store in data frame
    forest_data$Estimate[i] <- combined_effect
    forest_data$SE[i] <- sqrt(combined_variance)
}

# Create the forest plot with effect estimate on X-axis and PRS quantile on Y-axis
ggplot(forest_data, aes(x = Estimate, y = Quantile, xmin = Estimate - 1.96 * SE, xmax = Estimate + 1.96 * SE)) +
    geom_pointrange() +
    ylab("BMI.PRS SD") +
    xlab("Combined Effect Estimate") +
    theme_minimal()


```


## Let's do some 3d plots

```{r}
library(plotly)

# Assuming 'data' has 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
fig <- plot_ly(match_dat[match_dat$treat==1,], x = ~pretreat_bmi, y = ~delta, z = ~bmi.prs,
color = ~bmi.prs,
type = "scatter3d", mode = "markers")
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Baseline BMI'),
yaxis = list(title = 'Change in BMI'),
zaxis = list(title = 'PRS')),
title = "3D Plot of Change in BMI by PRS, Baseline BMI, and Treatment")
fig


# Assuming 'data' has 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
fig <- plot_ly(match_dat[match_dat$treat==0,], x = ~pretreat_bmi, y = ~delta, z = ~bmi.prs,
color = ~bmi.prs,
type = "scatter3d", mode = "markers")
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Baseline BMI'),
yaxis = list(title = 'Change in BMI'),
zaxis = list(title = 'PRS')),
title = "3D Plot of Change in BMI by PRS, Baseline BMI, and Treatment")
fig

```

# We examine distribution of delta by treatment and BMI

```{r}
load("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/merged_pheno_censor_final_withdrugs_smoke.rds")
pheno=dfh

match_dat=merge(match_dat,pheno,by.x="eid",by.y="identifier")

match_dat$enrollage=difftime(match_dat$f.53.0.0,match_dat$Birthdate,units = "days")/365.25
match_dat$first_bmi_age=difftime(match_dat$pretreat_measure_date,match_dat$Birthdate,units = "days")/365.25
library(table1)
dat=match_dat
dat$HyperLip_0_Any=factor(dat$HyperLip_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$Ht_0_Any=factor(dat$Ht_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$Dm_0_Any=factor(dat$Dm_0_Any,levels = c(1,2),labels = c("No","Yes"))
dat$f.31.0.0.x=factor(dat$f.31.0.0.x,levels = c(0,1),labels = c("Female","Male"))
dat$enrollage=as.numeric(dat$enrollage)
dat$smoke=factor(dat$smoke,levels=c(0,1),labels = c("No","Yes"))
dat$antihtn=factor(dat$antihtn,levels=c(0,1),labels = c("No","Yes"))
dat$statin=factor(dat$statin,levels=c(0,1),labels = c("No","Yes"))
dat$treat=factor(dat$treat,levels=c(0,1),labels=c("Untreated","Treated"))
label(dat$f.31.0.0.x)<- "Sex"
label(dat$enrollage) <- "Age"
label(dat$pretreat_bmi) <- "Pre-Treat BMI"
label(dat$HyperLip_0_Any) = "Hyperlipidemia"
label(dat$Ht_0_Any) = "Hypertension"
label(dat$Dm_0_Any) = "Diabetes Type 2"
dat$treat=as.factor(dat$treat)
label(dat$statin)="Statin"
label(dat$smoke)="Current Smoker"
label(dat$antihtn)="Anti Htn"
# One level of stratification
table1(~f.31.0.0.x + enrollage +pretreat_bmi+HyperLip_0_Any+Ht_0_Any+Dm_0_Any+statin+smoke+antihtn|treat,dat)

```

Now let's plot some stuff


```{r}
library(ggplot2)
        
ggplot(match_dat,aes(y=delta,x=treat,group=treat,fill=as.factor(treat)))+
  geom_boxplot()+labs(y="Change in Weight (kg) ",x="Treat",fill="Treat")+theme_classic()

ggplot(match_dat,aes(x=as.numeric(time_diff)/7,fill=as.factor(treat)))+geom_density()+
  labs(x="Weeks Between Measurements",main="Density",fill="Treat")+facet_wrap(~treat)+theme_classic()

pop.quants=quantile(prs_subset$BMI,probs = seq(0,1,by=0.25))
match_dat$prs.group=cut(match_dat$bmi.prs,breaks = pop.quants,labels = c(1:4))

levels(match_dat$treat) <- c("Untreated", "Treated")
ggplot(match_dat,aes(y=delta,x=prs.group,group=prs.group,fill=as.factor(prs.group)))+
  geom_boxplot()+labs(y="Change in Weight (kg)",x="PRS Quartile (population based)",fill="Prs Quartile")+theme_classic()+facet_wrap(~as.factor(treat),nrow = 1)



# Assuming you have a dataframe 'data' with columns 'baseline_BMI', 'change_in_BMI', and 'PRS'
match_dat$treat=factor(match_dat$treat,levels = c(0,1),labels=c("Untreated","Treated"))
ggplot(match_dat, aes(x = pretreat_bmi, y = delta,col=prs.group)) +
  geom_point() +  # Adjust the geom as necessary
  facet_grid(treat ~ prs.group) +
  labs(title = "Delta vs Baseline BMI by Treatment Status and PRS Quartiles",
       x = "Baseline BMI",
       y = "Delta BMI",
       col="BMI PRS Quartile",
       caption = "PRS Quartiles: 1 (Lowest) to 4 (Highest)") +
  theme_classic()

library(ggplot2)

# Assuming your dataframe is named 'data' with 'baseline_BMI', 'change_in_BMI', 'PRS', and 'treatment_group'
ggplot(match_dat, aes(x = pretreat_bmi, y = delta, size = bmi.prs, color = as.factor(treat))) +
  geom_point(alpha = 0.7) +  # Adjust alpha for bubble transparency if needed
  scale_size_continuous(range = c(1, 10)) +  # Adjust bubble size range as needed
  scale_color_brewer(palette = "Set1") +  # Choose a color palette that suits your data
  labs(title = "Bubble Plot of Change in BMI by Baseline BMI, PRS, and Treatment",
       x = "Baseline BMI",
       y = "Change in BMI",
       size = "PRS") +
  theme_minimal() +
  guides(color = guide_legend(title = "Treatment Group"), size = guide_legend(title = "PRS"))
```


And here we summarize by population quartiles:

```{r}
match_dat[match_dat$treat=="Treated",]%>%group_by(prs.group)%>%summarise(median(delta),IQR(delta),sd(delta))


match_dat[match_dat$treat=="Untreated",]%>%group_by(prs.group)%>%summarise(median(delta),IQR(delta),sd(delta))

dft=match_dat[match_dat$treat=="Treated",]%>%group_by(prs.group)%>%summarise(mean_change=median(delta),
                                               iqr=IQR(delta),
                                               sd=sd(delta),
                                              )
dfu=match_dat[match_dat$treat=="Untreated",]%>%group_by(prs.group)%>%summarise(mean_change=median(delta),
                                               iqr=IQR(delta),
                                               sd=sd(delta),
                                              )

library(ggplot2)


treat=ggplot(dft, aes(x = prs.group, y = mean_change)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_change-1.96*sd, ymax =  mean_change+1.96*sd), width = 0.1) +
  coord_flip() +  # Flips the axes
  xlab("PRS Quartile") +
  ylab("Change in BMI,Treated") +
  theme_minimal()

untreat=ggplot(dfu, aes(x = prs.group, y = mean_change)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_change-1.96*sd, ymax =  mean_change+1.96*sd), width = 0.1) +
  coord_flip() +  # Flips the axes
  xlab("PRS Quartile") +
  ylab("Change in BMI, Untreated") +
  theme_minimal()

ggarrange(untreat,treat,nrow=1)
```

```{r}
ggplot(data = match_dat,aes(x=bmi.prs,y=delta,col=as.factor(treat),fill=as.factor(treat)))+
  geom_point()+
  stat_smooth()+
  facet_wrap(~treat)+labs(y="Delta BMI",x="BMI PRS")+theme_classic()
```
