---
title: "glp1"
output: html_document
date: "2023-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

Here you can see the medication used according to NHS. This results in 2200 prescriptions for 940 people.

```{bash,eval=FALSE,echo=TRUE}
#in bash
grep -E 'Byetta|Ozempic|Wegovy|Bydureon|Trulicity|Lixumia|Victoza|Rybelsus|Liraglutide|Dulaglutide|Exenatide|Lixisenatide|Semaglutide' gp_scripts.txt > glp1s_withi.txt
```


We match on sex, DOB, and BMI at baseline and display characteristics of population before and after matching.
```{r}
source("~/multistate2/code/utils.R")
prs_subset=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/prs_subset.rds")
glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s_withi.txt")
colnames(glp1)[1]=c("Identifier")
colnames(glp1)[3]=c("event_dt")
df_baseline=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/dfukb_baseline.rds")
## treatment data 
glp1$event_dt <- as.Date(glp1$event_dt, format="%d/%m/%Y")
unique_patients <- glp1 %>%
  group_by(Identifier) %>%
  summarize(earliest_event_dt = min(event_dt))

m=merge(prs_subset,unique_patients,by = "Identifier",all.x = T)
m2=merge(m,df_baseline,by.x = "Identifier",by.y="identifier")
m2=m2[!is.na(m2$f.21001.0.0),]
m2$treat=as.integer(ifelse(!is.na(m2$earliest_event_dt),1,0))


library(MatchIt)
````


Now BMI data:

```{r}
biomarkers=readRDS("~/Library/CloudStorage/Dropbox-Personal/ukbb-ehr-data/data/biomarkers.rds")
bmi=biomarkers[biomarkers$variable%in%"bmi",]
bmi$date <- as.Date(bmi$date, format="%Y/%m/%d")
# sorted_data <- bmi %>%
#   arrange(eid, date) %>%
#   group_by(eid)  %>%
#   slice(n())
# 
# 
# mg=merge(m2,sorted_data,by.x="Identifier",by.y="eid")
# mg$bmi.prs=mg$BMI

df_baseline=df_baseline[!is.na(df_baseline$f.21001.0.0),]
baseline_bmi=data.frame("eid"=df_baseline$identifier,"date"=as.Date(df_baseline$f.53.0.0),"variable"="bmi",source="ukb","value"=df_baseline$f.21001.0.0)
bmi2=rbind(bmi,baseline_bmi)

## merge with the date of treatment for treated folks
grc=bmi2[!(bmi2$eid%in%unique_patients$Identifier),]
grc$bmi_date=as.Date(grc$date)


## treated
data2=merge(bmi2,unique_patients,by.x="eid",by.y="Identifier")
data2$bmi_date=as.Date(data2$date)
data2$script_date=as.Date(data2$earliest_event_dt)



last_bmi_before_script2 <- data2 %>%
  group_by(eid) %>%
  filter(bmi_date < script_date) %>%
  slice_max(order_by = bmi_date, n = 1) %>%
  ungroup()

lbmi=last_bmi_before_script2 %>%
               distinct(eid, .keep_all = TRUE)

lbmi$diff=difftime(lbmi$script_date,lbmi$bmi_date,units = "days")

last_bmi_after_script <- data %>%
  group_by(eid) %>%
  filter(bmi_date > script_date) %>%
  slice_max(order_by = bmi_date, n = 1) %>%
  ungroup()

bmis_treated=merge(lbmi,last_bmi_after_script[,c("eid","bmi_date","value")],by="eid")
bmis_treated$delta=bmis_treated$value.y-bmis_treated$value.x
bmis_treated$duration=difftime(bmis_treated$bmi_date.y,bmis_treated$script_date,units = "days")
#only keep those with bmis measured at least 90 days out
bmis_treated=bmis_treated[bmis_treated$duration>90,]


## for controls first bmi will be baseline

controls_last <- grc %>%
  group_by(eid) %>%
  slice_max(order_by = bmi_date, n = 1) %>%
  ungroup()

controls_last=controls_last %>%
               distinct(eid, .keep_all = TRUE)
controls_all=merge(controls_last,df_baseline,by.x = "eid",by.y="identifier")

controls_all$delta=controls_all$value-controls_all$f.21001.0.0
controls_all$duration=difftime(controls_all$date,controls_all$f.53.0.0,units = "days")
## Now merge

for_matching_treat=data.frame("eid"=bmis_treated$eid,"firstval"=bmis_treated$value.x,"treatdate"=as.Date(bmis_treated$script_date),"lastval"=bmis_treated$value.y,"duration"=bmis_treated$duration,"delta"=bmis_treated$delta,"treat"=1)

for_matching_controls=data.frame("eid"=controls_all$eid,"firstval"=controls_all$f.21001.0.0,
                                 "treatdate"=controls_all$f.53.0.0,"lastval"=controls_all$value,
                                 "duration"=controls_all$duration,
                                 "delta"=controls_all$value-controls_all$f.21001.0.0,"treat"=0)


for_matching=rbind(for_matching_treat,for_matching_controls)

for_matching=merge(for_matching,df_baseline,by.x = "eid",by.y="identifier")

## only keep those whose last bmi measurement is after treatment and matchon this 

matched=matchit(treat~f.34.0.0+f.21001.0.0+f.31.0.0,data=for_matching)
summary(matched)
datm=match.data(matched)

datm=merge(datm,prs_subset,by.x="eid",by.y="Identifier")
## grab the control delta as difference between the baseline BMI and the last measured BMI
datm$bmi.prs=datm$BMI
```

## Within treated populiaton

```{r}

datm$delta=(datm$value-datm$f.21001.0.0)/datm$f.21001.0.0

your_model=lm(delta~bmi.prs+f.31.0.0,dat=datm[datm$treat==1&datm$f.21001.0.0>30,])

```



We now measure the percentage of BMI chance and calculate the influence of treatment, PRS, or the interaction:

```{r}


your_model=lm(scale(delta)~treat+treat:bmi.prs+f.31.0.0+f.21001.0.0,dat=datm)
your_model=lm(scale(delta)~treat+treat:bmi.prs,dat=datm,weights = weights)



model_summary <- summary(your_model)

model_summary

coefficients <- model_summary$coefficients
rownames(coefficients)=c("(Intercept)","Treatment","Treatment:PRS")
# Create a data frame of PRS values (e.g., from min to max PRS in your data)
prs_values <- data.frame(PRS = seq(min(mg$BMI), max(mg$BMI), length.out = 100))

# Calculate predicted treatment effect for each PRS value
# Adjust the formula based on your model's structure
prs_values$Predicted_Effect <- coefficients["(Intercept)", "Estimate"] +
  coefficients["Treatment", "Estimate"] * 1 +
  coefficients["Treatment:PRS", "Estimate"] * 1 * prs_values$PRS

# Plot
library(ggplot2)
ggplot(prs_values, aes(x = PRS, y = Predicted_Effect)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Predicted Treatment Effect Across PRS",
       x = "Polygenic Risk Score (PRS)",
       y = "Predicted Treatment Effect")

```


# perhaps we should control for other risk conditions
* let's also match for Hyperlipidemia, Dm, and CAD

```{r}

pheno=readRDS("~/Library/CloudStorage/Dropbox-Personal/df_ukb_pheno_updated.rds")

mgmas=merge(mg,pheno,by="Identifier")
mgmas$enrollage=difftime(mgmas$Enrollment_Date,mgmas$Birthdate.x,units = "days")/365.25

mgmas=mgmas[!is.na(mgmas$f.34.0.0)&!is.na(mgmas$f.21001.0.0)&!is.na(mgmas$f.31.0.0),]
mgmas$CAD=scale(mgmas$CAD)
matched=matchit(treat~f.34.0.0+f.21001.0.0+f.31.0.0+DmT2_Any+HyperLip_Any+Ht_Any+Cad_Any,data=mgmas)
```


# extract matched data

* so now we've matched on comorbidities, and we'll extract the effect of treatment, or treatment + interaction

```{r,echo=T}
summary(matched)
datm=match.data(matched)

datm$delta=(datm$value-datm$f.21001.0.0)/datm$f.21001.0.0

your_model=lm(scale(delta)~treat+treat:bmi.prs,dat=datm)

your_model=lm(scale(delta)~treat,dat=datm)
model_summary <- summary(your_model)

model_summary
```

## but now when we add back in interaction fo bmi on treatment to this comorbid matched population
```{r,echo=T}
your_model=lm(scale(delta)~treat+treat:bmi.prs,dat=datm)
model_summary <- summary(your_model)

model_summary
```

## Now let's try CAD PRS and matching for known CAD risk factors

* Here we match for Hyperlipidemia, Htn, and Dm
* We exclude individuals with CAD at enrollment *or* prior to beginning drug.


```{r}

w=which(!is.na(mg$earliest_event_dt))
mg$age.script=1000
mg$age.script[w]=difftime(as.Date(mg$earliest_event_dt[w]),time2 = as.Date(mg$Birthdate[w]),units = "days")/365.25

pheno=readRDS("~/Library/CloudStorage/Dropbox-Personal/df_ukb_pheno_updated.rds")

mgmas=merge(mg,pheno,by="Identifier")

mgmas$enrollage=difftime(mgmas$Enrollment_Date,mgmas$Birthdate.x,units = "days")/365.25

#eliminate those with CAD at baseline
mgmas=mgmas[mgmas$Cad_censor_age>mgmas$enrollage,]
## remove those with CAD prior to script date
mgmas=mgmas[-which(mgmas$age.script<mgmas$Cad_censor_age&mgmas$Cad_Any==2),]

mgmas=mgmas[!is.na(mgmas$f.34.0.0)&!is.na(mgmas$f.21001.0.0)&!is.na(mgmas$f.31.0.0),]
mgmas$CAD=scale(mgmas$CAD)
matched=matchit(treat~f.34.0.0+f.21001.0.0+f.31.0.0+DmT2_Any+HyperLip_Any+Ht_Any,data=mgmas)
summary(matched)
datm=match.data(matched)

datm$cadevent=ifelse(datm$Cad_Any==2,1,0)

datm$enrollage=as.numeric(datm$enrollage)

datm$cad.prs=datm$CAD
```

We need to create a time dependent variable for drug start time so we don't unfairly bias for people who survived until treatment as being in the treated arm during those 'immortal' years;

```{r, eval=T, echo=T}
data_period_1 <- datm[datm$treat==1,] %>%
  mutate(start = enrollage, stop = age.script, treatment = 0)

data_period_2 <- datm[datm$treat==1,] %>%
  mutate(start = age.script, stop = Cad_censor_age, treatment = 1)

# Combining the data back together
data_long <- bind_rows(data_period_1, data_period_2)

data_period_1un <- datm[datm$treat==0,] %>%
  mutate(start = enrollage, stop = Cad_censor_age, treatment = 0)

# Combining the data back together
data_long <- bind_rows(data_period_1un, data_long)


```

* Here we test for the effect of drug, CAD prs, and drug on cad PRS, using age as time scale:

```{r}
coxph(Surv(time=enrollage,time2=Cad_censor_age,event=cadevent)~treat+cad.prs+treat:cad.prs,data = data_long)
```

So it appears that while the CAD PRS does increase risk for CAD (appropriately) there is no differential effect of treatment on CAD response.


