---
title: "MoreSelective"
output: html_document
date: "2023-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

## let's be more selective and select first and last script date

* Here we grab the first script date and the last script date for treated individuals

```{r}
source("~/multistate2/code/utils.R")
library(MatchIt)
prs_subset=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/prs_subset.rds")
#glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s.txt")
df_baseline=readRDS("~/Library/CloudStorage/Dropbox-Personal/pheno_dir/output/dfukb_baseline.rds")

glp1=fread("~/Library/CloudStorage/Dropbox-Personal//glp1s_withi.txt")
colnames(glp1)[1]=c("Identifier")
colnames(glp1)[3]=c("event_dt")
glp1$event_dt <- as.Date(glp1$event_dt, format="%d/%m/%Y")


biomarkers=readRDS("~/Library/CloudStorage/Dropbox-Personal/ukbb-ehr-data/data/biomarkers.rds")
bmi=biomarkers[biomarkers$variable%in%"bmi",]
bmi$date <- as.Date(bmi$date, format="%Y/%m/%d")


patient_med_data <- glp1 %>%
  group_by(Identifier) %>%
  summarize(first_script_date = min(event_dt),last_script_date = max(event_dt))

colnames(patient_med_data)=c("eid","first_script_date","last_script_date")
patient_med_data$first_script_date <- as.Date(patient_med_data$first_script_date, format = "%Y-%m-%d")
patient_med_data$last_script_date <- as.Date(patient_med_data$last_script_date, format = "%Y-%m-%d")
```

## Treated Folks

* We grab the first BMI as the median BMI that is within 1 year of the first script date
* We grab the post-treat BMI as the median BMI that is within 12 weeks after 1st script to 1 year after last.
* We tried the last measurement before but it is noisy to use the final measurement only, we can also see how many measurements folks had

```{r}
patient_bmi_data=bmi[bmi$eid%in%patient_med_data$eid,c("eid","date","value")]
bbmi=df_baseline[,c("identifier","f.53.0.0","f.21001.0.0")]
bbmi$f.53.0.0=as.Date(bbmi$f.53.0.0)

patient_bmi_data=rbind(patient_bmi_data,bbmi[,c("identifier","f.53.0.0","f.21001.0.0")],use.names=FALSE)

patient_bmi_data$date <- as.Date(patient_bmi_data$date, format = "%Y-%m-%d")
merged_data <- merge(patient_bmi_data, patient_med_data, by = "eid")

bmi_before_script <- merged_data %>%
  filter(date < first_script_date & date > (first_script_date - years(1)))%>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"medbmi"=median(value),"number"=length(value))

summary(bmi_before_script$number)

bmi_after_script <- merged_data %>%
  filter(date > first_script_date + weeks(12) & date < last_script_date + years(1)) %>%
  
  group_by(eid) %>%
  summarise("afterdate"=min(date),"medbmi"=median(value),"number"=length(value))

summary(bmi_after_script$number)
```

## Now we merge the first and last BMI measurements with the script data and perform checks
* to make sure first measure is before drug start
*  check to make sure last BMI measure is at least 12 weeks after drug start

```{r}
final_data <- (bmi_before_script[,c("eid","beforedate","medbmi")]) %>%
  left_join(patient_med_data[,c("eid","first_script_date","last_script_date")], by = "eid") %>%
  left_join(bmi_after_script[,c("eid","afterdate","medbmi")], by = "eid")
colnames(final_data)=c("eid","pretreat_measure_date","pretreat_bmi","first_script_date","last_script_date","post_drug_measure_date","post_drug_bmi")

final_data=na.omit(final_data)

summary(year(final_data$pretreat_measure_date))
summary(year(final_data$post_drug_measure_date))

# check to make sure first measure is before drug start
sum(difftime(final_data$first_script_date,final_data$pretreat_measure_date)<0)

# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data$post_drug_measure_date,final_data$first_script_date,units = "days")<83)
## controls

```

Less clearcut for controls
# we will choose controls to mirror our selection above
* First BMI median between the treated populations pretreat measure dates (min,max pre_treat_date)
* Last BMI median between the treated populations post-treat measure dates

```{r}
controls=bmi[!bmi$eid%in%patient_med_data$eid,]

## since there is no med, grab first and last measruement 
## find measruements around the time of treated

bmi_controls_first <- controls %>%
  filter(date >= min((final_data$pretreat_measure_date)) & date < max((final_data$pretreat_measure_date))) %>%
  group_by(eid) %>%
  summarise("beforedate"=min(date),"meanbmi"=mean(value),"number"=length(value))


summary(bmi_controls_first$beforedate)
summary(bmi_controls_first$number)

bmi_controls_last <- controls %>%
  filter(date >= min(final_data$post_drug_measure_date) & date < max(final_data$post_drug_measure_date))  %>%
  group_by(eid) %>%
 summarise("afterdate"=min(date),"meanbmi"=mean(value),"number"=length(value))

summary(bmi_controls_last$afterdate)
summary(bmi_controls_last$number)

final_data_controls=
  inner_join(bmi_controls_first[,c("eid","beforedate","meanbmi")], bmi_controls_last[,c("eid","afterdate","meanbmi")],by = "eid", suffix = c("_before", "_after"))

# remote those in which dates are the same

final_data_controls=final_data_controls[final_data_controls$beforedate!=final_data_controls$afterdate,]
```

Again, checks:


```{r}
# check to make sure last BMI measure is at least 12 weeks after drug start
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")>0)
sum(difftime(final_data_controls$afterdate,final_data_controls$beforedate,units = "days")==0)

## controls
summary(year(final_data_controls$beforedate))
summary(year(final_data_controls$afterdate))

final_data$treat=1
final_data_controls$treat=0

treat=final_data[,c("eid","pretreat_bmi","post_drug_bmi","pretreat_measure_date","post_drug_measure_date","treat")]
untreat=final_data_controls[,c("eid","meanbmi_before","meanbmi_after","beforedate","afterdate","treat")]
names(untreat)=names(treat)
all_data=rbind(treat,untreat)
all_data$time_diff=difftime(all_data$post_drug_measure_date,all_data$pretreat_measure_date,units = "days")

df=merge(all_data,df_baseline[,c("identifier","f.34.0.0","f.31.0.0")],by.x="eid",by.y="identifier")
df_with_prs=merge(df,prs_subset[,c("Identifier","BMI")],by.x="eid",by.y="Identifier")
```

## now we perform matching on

* duration between measurements
* patient birthday
* sex
* pretreat BMI

```{r}
## now match on pretreat bmi
matcher=matchit(treat~pretreat_bmi+f.34.0.0+f.31.0.0+time_diff,data = df_with_prs)
summary(matcher)

#extract matched data
match_dat=match.data(matcher)
match_dat$bmi.prs=match_dat$BMI
match_dat$delta=match_dat$post_drug_bmi-match_dat$pretreat_bmi
summary(lm(delta~treat+treat:bmi.prs,match_dat))
```


# We examine distribution of delta by treatment and BMI

```{r}
boxplot(delta~treat,data = match_dat,ylab="Change in BMI")
hist(as.numeric(match_dat$time_diff[match_dat$treat==1]/7),xlab="Weeks on treatment",main="Density")

match_dat$prs.group=cut(match_dat$bmi.prs,breaks = quantile(match_dat$bmi.prs,probs = seq(0,1,by=0.25)),labels = c(1:4))
boxplot(delta~prs.group,data = match_dat,ylab="Change in BMI",xlab="BMI.PRS Quartile")
```

