---
title: "saturn"
output: html_document
date: "2023-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r livesaved}
modelfit=fitfunc2(data.table(train),ages = ages,nstates = nstates,mode = "binomial",covariates ="cad.prs+f.31.0.0+smoke+antihtn_now+statin_now")

b=coefplotsmooth2(ages = ages,start = "Health",stop = "Cad",modelfit = modelfit,window_width = 20,span = 0.75,degree = 2)
coefs=b$custom_smooth

prs_quants=qnorm(c(seq(0.1,0.9,by=0.1)))

prs_quants=c(-2:2)

sex=c(0,1)

atrisk=expand.grid(intercept=1,prs_quants=prs_quants,sex=sex,smoke=c(0,1),ah=c(0,1),stat=c(0,1))

ar=atrisk[atrisk$ah==0&atrisk$smoke==0&atrisk$stat==0,]
agesint=seq(40,70,by=5)


tenlifarray=array(NA,dim=c(length(agesint),nrow(ar),4))
dimnames(tenlifarray)=list(agesint,levels(interaction(prs_quants,sex)),c("Ten","Lifetime","TenBenefit","LifetimeBenefit"))


for(g in 1:length(agesint)){
  age=agesint[g]
  
  tenlifarray[g,,1]=compute_prediction_product_matrix(atrisk = ar,agepredinterval = age:(age+10),coefmat = coefs,RR = 0.80)$PredictedIntervalrisk

  tenlifarray[g,,2]=compute_prediction_product_matrix(atrisk = ar,agepredinterval = age:80,coefmat = coefs,RR = 0.80)$PredictedIntervalrisk
  
  tenlifarray[g,,3]=compute_prediction_product_matrix(atrisk = ar,agepredinterval = age:(age+10),coefmat = coefs,RR = 0.80)$Hazard_treated
  
  tenlifarray[g,,4]=compute_prediction_product_matrix(atrisk = ar,agepredinterval = age:80,coefmat = coefs,RR = 0.80)$Hazard_treated
  
}

netben=tenlifarray[,,"Lifetime"]-tenlifarray[,,"LifetimeBenefit"]
ben=melt(netben)
ben2=melt(tenlifarray[,c(6:10),"LifetimeBenefit"])
ten=melt(tenlifarray[,c(6:10),"Ten"])
l=melt(tenlifarray[,c(6:10),"Lifetime"])

mer=merge(ten,l,by=c("Var1","Var2"))
mer=merge(mer,ben,by=c("Var1","Var2"))

# colnames(mer)=c("Age","PRS Category","Ten Year","Lifetime","Net Ben")
# mer$`PRS Category`=factor(mer$`PRS Category`,labels = c(0.01,seq(0.1,0.9,by=0.1),0.99))
# mermelt=melt(mer,id.vars=c("Age","PRS Category"))

colnames(mer)=c("Age","PRS Category","Ten Year","Lifetime","Net Ben")
mer$`PRS Category`=factor(mer$`PRS Category`,labels = -2:2)
mermelt=melt(mer,id.vars=c("Age","PRS Category"))


## dot point

g1=ggplot(mer,aes(`Ten Year`,y=`Lifetime`,ymin = `Lifetime`-0.001,ymax=`Lifetime`+0.001,color=`PRS Category`,group=`Age`))+geom_point()+geom_pointrange()+geom_line()+theme_classic(base_size = 20)+labs(x="MS Predicted Ten Year Risk",y="MS Lifetime")


g1=ggplot(mer,aes(`Ten Year`,y=`Lifetime`,ymin = `Lifetime`-0.01,ymax=`Lifetime`+0.01,color=`PRS Category`,group=`Age`))+geom_point()+geom_pointrange()+geom_line()+theme_classic(base_size = 20)+labs(x="MS Predicted Ten Year Risk",y="MS Lifetime")


ggsave(plot = g1,filename = "../output/MStenvsMslifesaturn.pdf",dpi=300,height = 4,width = 6)

g2=ggplot(mer,aes(`Ten Year`,y = `Net Ben`,ymin=`Net Ben` -0.001,ymax=`Net Ben`+0.001,color=`PRS Category`,group=`Age`))+geom_point()+geom_pointrange()+geom_line()+theme_classic(base_size = 20)+labs(x="MS Predicted Ten Year Risk",y="MS Lifetime Benefit",color="PRS Percentile")

ggsave(plot = g2,filename = "../output/MStenvsMslifebenefit.pdf",dpi=300,height = 4,width = 6)

ggsave(
ggarrange(g1,g2,benefit,ncol=2,nrow=2,common.legend = T),file="../output/saturnbenefit_better.pdf",dpi=600,height = 12,width=10)
