---
title: "topictry"
output: html_document
date: "2023-12-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

## Overview of Topic Modeling

We first lay out a basic topic model without any accounting for time. 

In this model, we have a number of people, M.

Each individual has a number of diagnoses, N_{i=1\dots M}.We will simulate this as a poisson RV with mean of 10 diagnoses per patient.

There are a total number of $N=\sum _{i=1}^{M}N_{i}$ diagnoses over all individuals.

There are a total of $V$ disease codes entirely.

For each person, they have a distribution of topic responsibility on K topics, $\theta_{i}$. 

This topic distribution, $\theta_i$ will be sparsely simulated so individuals are loaded on a minimal number of topics using a random dirichlet with $\alpha$ = 0.1_{RK}.

For each of the K topics, there is a distribution on diseases. There are a total of V diseases.
This parameter $\phi$, over diseases, is simulated also according to a sparse dirichlet, this time with hyperparameter $\beta$ for example also = 0.1_RV

Now each disease arises from a topic (there can be overlap such that one disease can be featured in several topics) with latent indicator $z_{i=1..M,j=1:Ni}$ with an integer from 1:K. $1:Ni$ here represents the fact that an individual can have up to Ni diagnoses, so that $Z$ is a giant vector of length N.

Now according to the assigned topic (1:K) for that diagnosis, the probability of the given diagnosis is drawn according to $\phi_{kd}$


```{r cars}
library(rBeta2009)
library(ggplot2)
library(reshape2)
library(splines)

# number of people
M=100

# number of topics
K=5

## number of diagnoses per patient, mean of 10 diagnoses per patient
Nd=rpois(M,lambda = 10)

## total numer of diagnoses
N=sum(Nd)

# number of disease total
V=1000
```


```{r pressure, echo=FALSE}
## make sparse so that most mass concentrated on one topic 


alpha=rep(0.1,K)

## make sparse so that most mass concentrated on one word 
beta=rep(0.1,V)

## distribution of topics per pt
theta=rdirichlet(M,shape = alpha)
dim(theta)

## distribution of words per topic
phi=rdirichlet(K,shape = beta)
dim(phi)

disease <- list() # Initialize the list outside the loop
for(i in 1:M){
  disease[[i]] <- vector("list", Nd[i]) # Initialize the ith list
  for(j in 1:Nd[i]){
    zij <- which(rmultinom(1, size = 1, prob = theta[i,]) == 1)
    wij <- which(rmultinom(1, 1, prob = phi[zij,]) == 1)
    disease[[i]][j] <- wij # diagnose your patient with some badness 
  }
}

```

Show the distirbtuion of theta and phi

```{r}

ggplot(data = melt(theta),aes(x=value,fill=value,group=value))+geom_histogram()+labs(x="Theta_m",y="Density")+
  theme_classic()

ggplot(data = melt(phi),aes(x=value,fill=value,group=value))+geom_histogram()+labs(x="Phi_k",y="Density")+theme_classic()
```


```{r}
# Assuming 'disease' is your list of words for each person
# Convert 'disease' list into a data frame for plotting
disease_data <- data.frame(person = rep(seq_along(disease), sapply(disease, length)),
                        disease = unlist(disease))

# Now create the plot
random_peeps=sample(M,size = 5)
ggplot(disease_data[disease_data$person%in%random_peeps,], aes(x = disease,fill=as.factor(disease))) +
  geom_bar() +
  facet_wrap(~ person, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Disease", y = "Frequency", title = "Disease Distribution per Person",fill="Disease")+theme_classic()
```

## Adding a time element

Now, in McVean et al, instead of having only one phi_k per topic, there are phi_kt for a time and topic specific diagnosis distribution.

Easy enough! Recall before we had a matrix that was KxV for the disease distribtuion per topic.

Now we will initialize an *array* that is KxVxT that has a disease distribution per topic per time. 

The trick is that these need to be somewhat continuous -- you can't go from a 100% probability of hypertension to 0 overnight. So ... they use a natural spline.


```{r}
K <- length(alpha)  # Number of topics
V <- length(beta)   # Number of diagnoses
T <- 10             # Number of time points
time_points <- seq(0, 10, length.out = T)
phi <- array(dim = c(K, V, T))

# Generate the spline for each topic and diagnosis
for (k in 1:K) {
  for (v in 1:V) {
    # Generate coefficients for the spline - should be of length 5
   
    # Generate a B-spline basis
    knots <- c(0.5) # Example knot
    degree=3
    bspline_basis <- bs(time_points, knots = knots, degree = degree, intercept = TRUE)
    coef <- runif(length(knots)+degree+1) # Adjust based on the number of columns in the bspline_basis


    # Calculate the spline values
    spline_values <- bspline_basis %*% coef

    # Store the values in the phi array
    phi[k, v, ] <- spline_values
  }
}

# Normalize probabilities in phi
phi <- apply(phi, c(1, 3), function(x) {
  x <- pmax(x, 0) # Ensure non-negative
  x / sum(x)      # Normalize
})

dim(phi)
```

Now $\phi$ is the array of diseases x topics x time, so for a given topic, this should be continous. Let's look at a sample topic with a few diseases to check:

```{r}
# Assuming phi is already created and is a KxVxT array
# Choose a topic to plot, for example, the first topic
selected_topic <- 2
selected_diseases=sample(V,size = 10)

# Extract the data for this topic
topic_data <- phi[c(selected_diseases),selected_topic, ]

# Create a sequence of time points for the x-axis
time_points <- seq(0, 10, length.out = dim(phi)[3])

# Plotting
matplot(time_points, t(topic_data), type = 'l', lty = 1,
        xlab = "Time", ylab = "Probability",
        main = paste("Disease Probabilities Over Time for Topic", selected_topic))
legend("topright", legend = paste("Disease", 1:dim(phi)[2]), col = 1:dim(phi)[2], lty = 1)
```


## However this assumes that an individual stays within a topic

* Assumption: that the change in disease distribtuion for a given topic captures all the heterogeneity of disease coding
* We know that is not the case from our previous work: an individual can switch to a new pattern based on genetics and environmental factors, and we should have an ability to 'update' an individuals' topic distribtuion

* Furthermore, genetics has a *declining* importance over time, and the new diagnoses carry more weight. The rate of advancement *through a topic* (i.e., the shape of the spline) also may be affected by these intrinsic factors

* Finally, occupancy in one topic necessarily may prevent other topics

# Example function to update theta based on new diagnosis

Let's start by modeling an individual's transition to a new topic distribtuion as a funciton of genetics:

```{r}
update_theta <- function(current_theta, new_diagnosis, phi_time_dependent, genetics_influence, current_time, max_time) {
  # Decay factor for genetics influence
  decay_factor <- (max_time - current_time) / max_time
  adjusted_genetics_influence <- genetics_influence * decay_factor

  # Select phi for the current time
  phi_current_time <- phi_time_dependent[,,current_time]

  # Adjust prior based on genetics
  adjusted_prior <- current_theta * adjusted_genetics_influence

  # Calculate the likelihood of new diagnosis for each topic
  likelihood <- phi_current_time[new_diagnosis,] / colSums(phi_current_time)

  # Bayesian update
  updated_theta <- likelihood * adjusted_prior
  return(updated_theta / sum(updated_theta))  # Normalize the updated theta
}

M <- 100          # Number of patients
K <- 5            # Number of topics
V <- 1000           # Number of diseases
T <- 10           # Number of time steps
max_time <- T     # Assuming T is the maximum time

# Initial topic distributions (theta) for each patient
initial_theta <- rdirichlet(M, shape = rep(0.1, K))

# Time-varying disease distributions within topics (phi)
# Assuming phi is a 3D array (topics x diseases x time)
# This needs to be defined based on your model

# Genetics influence (for simplicity, assuming a constant vector for each patient)
genetics_influence <- matrix(runif(M * K, 0.5, 1.5), nrow = M, ncol = K)

# Structure to hold theta values over time for each patient
theta_over_time <- array(dim = c(T, M, K))
theta_over_time[1, , ] <- initial_theta

# Function to simulate a new diagnosis (placeholder)
simulate_new_diagnosis <- function() {
  # Randomly simulate a new diagnosis
  return(sample(1:V, 1))
}

for (time in 2:T) {
  for (patient in 1:M) {
    # Simulate a new diagnosis for this patient at this time
    new_diagnosis <- simulate_new_diagnosis()

    # Update the patient's topic distribution based on the new diagnosis
    theta_over_time[time, patient, ] <- update_theta(
      current_theta = theta_over_time[time - 1, patient, ],
      new_diagnosis = new_diagnosis,
      phi_time_dependent = phi,  # Assuming phi is defined
      genetics_influence = genetics_influence[patient, ],
      current_time = time,
      max_time = max_time
    )
  }
}

library(ggplot2)
library(reshape2)

# Select data for a specific patient, e.g., patient 1
pt=sample(M,size = 1)
theta_patient1 <- theta_over_time[,pt , ]

# Convert to a data frame and reshape
theta_patient1_df <- as.data.frame(theta_patient1)
colnames(theta_patient1_df) <- paste("Topic", 1:K, sep = "")
theta_patient1_df$Time <- 1:nrow(theta_patient1_df)

theta_patient1_long <- melt(theta_patient1_df, id.vars = "Time", variable.name = "Topic", value.name = "Proportion")

# Plot
ggplot(theta_patient1_long, aes(x = Time, y = Proportion, fill = Topic)) +
  geom_area(position = 'stack') +
  labs(x = "Time", y = "Topic Proportion", fill = "Topic") +
  ggtitle("Distribution of Topics for Patient 1 Over Time")
```

## TODO: affect speed based through transitions based on genetics
```{r}
# Pseudocode - Adjusting spline coefficients based on a regression model

# Assume a regression model is built to predict coefficients from genetic data
# coef_model <- lm(coef ~ genetic_factors, data = training_data)

# For each individual
for (i in 1:N_individuals) {
  # Get individual's genetic data
  genetic_data_i <- get_genetic_data_for_individual(i)

  # Predict/adjust coefficients based on genetic data
  predicted_coef <- predict(coef_model, newdata = genetic_data_i)

  # Use these coefficients for the spline
  bspline_basis <- bs(time_points, knots = knots, degree = degree, intercept = TRUE)
  spline_values <- bspline_basis %*% predicted_coef
  phi_individual[i, , ] <- spline_values
}
```