---
title: "topictry"
output: html_document
date: "2023-12-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```

## Overview of Topic Modeling

We first lay out a basic topic model without any accounting for time. 

\item In this model, we have a number of people, M.

\item Each individual has a number of diagnoses, N_{i=1\dots M}.We will simulate this as a poisson RV with mean \lambda of 10 diagnoses per patient.

\item There are a total number of $N=\sum _{i=1}^{M}N_{i}$ diagnoses over all individuals.

\item There are a total of $V$ disease codes entirely.

\item For each of the M individuals, there exists a distribution over K topics, $\theta_{i}$. 

\item This topic distribution for a given individual, $\theta_i$ will be sparsely simulated so individuals are loaded on a minimal number of topics using a random dirichlet with $\alpha$ = [0.1,...K] .

\item For each of the K topics, there is a distribution on V diseases. There are a total of V diseases.
This parameter $\phi_k$, over diseases, is simulated also according to a sparse dirichlet, this time with hyperparameter $\beta$ for example also = 0.1_RV

\item Here we allow that each topic has a special 'topic specific' disease enrichment, such that for a select number d of randomly sampled diseases, here d=10, \beta is (V-d)/V with the remaining \beta = 1/V.

# Generative model 
Now each disease arises from a topic (there can be overlap such that one disease can be featured in several topics) with latent indicator $z_{i=1..M,j=1:Ni}$ with an integer from 1:K.

$1:Ni$ here represents the fact that an individual can have up to Ni diagnoses, so that $Z$ is a giant vector of length N.

According to the assigned topic (1:K) for that diagnosis, the probability of the given diagnosis is drawn according to $\phi_{kd}$



# Table of Variables

Variable | Type | Meaning
--- | --- | ---
K | integer | number of topics (e.g. 50)
V | integer | number of diseases in the vocabulary (e.g. 50,000 or 1,000,000)
M | integer | number of persons
\(N_{i_{1..M}}\) | integer | number of diseases in person \(i\)
N | integer | total number of diseases in all persons; sum of all \(N_d\) values, i.e. \(N = \sum_{i=1}^{M} N_i\)
\(\alpha_{k_{1..K}}\) | positive real | prior weight of topic \(k\) in a person; usually the same for all topics; normally a number less than 1, e.g. 0.1, to prefer sparse topic distributions, i.e. few topics per person
\(\alpha\) | K-dimensional vector of positive reals | collection of all \(\alpha_k\) values, viewed as a single vector
\(\beta_{w_{1..V}}\) | positive real | prior weight of disease \(w\) in a topic; usually the same for all diseases; normally a number much less than 1, e.g. 0.001, to strongly prefer sparse disease distributions, i.e. few diseases per topic
\(\beta\) | V-dimensional vector of positive reals | collection of all \(\beta_w\) values, viewed as a single vector
\(\phi_{k_{1..K},w_{1..V}}\) | probability (real number between 0 and 1) | probability of disease \(w\) occurring in topic \(k\)
\(\Phi_{k_{1..K}}\) | V-dimensional vector of probabilities, which must sum to 1 | distribution of diseases in topic \(k\)
\(\theta_{d_{1..M},k_{1..K}}\) | probability (real number between 0 and 1) | probability of topic \(k\) occurring in person \(d\)
\(\Theta_{d_{1..M}}\) | K-dimensional vector of probabilities, which must sum to 1 | distribution of topics in document \(d\)
\(z_{d_{1..M},w_{1..N_d}}\) | integer between 1 and K | identity of topic of disease \(w\) in document \(d\)
Z | N-dimensional vector of integers between 1 and K | identity of topic of all diseases in all documents
\(w_{d_{1..M},w_{1..N_d}}\) | integer between 1 and V | identity of disease \(w\) in person \(d\)
W | N-dimensional vector of integers between 1 and V | identity of all diseases in all persons

## Simulation


```{r cars}
set.seed(123)
library(rBeta2009)
library(ggplot2)
library(reshape2)
library(splines)

# number of people
M=100

# number of topics
K=5

## number of diagnoses per patient, mean of 10 diagnoses per patient
Nd=rpois(M,lambda = 10)

## total numer of diagnoses
N=sum(Nd)

# number of disease total
V=300
```


```{r sparse, echo=TRUE}
## make sparse so that most mass concentrated on one topic 
alpha=rep(0.1,K)

## make sparse so that most mass concentrated on one disease 
beta=rep(1/V,V)

## distribution of topics per pt
theta=rdirichlet(M,shape = alpha)
dim(theta)

## distribution of diseases per topic, load heavily on some
# d will be number of strong disease per topic

d=10
topic_spec_disease=matrix(sample(V,size = K*d),nrow=K,byrow = T)
phi=rdirichlet(K,shape = beta)

for (k in 1:K) {
    for (v in 1:V) {
        # Assign higher probabilities to certain diseases for each topic
        # Placeholder: Assigning higher weights to some diseases
        if (v%in%topic_spec_disease[k,]) {
            beta[v] <- (V-d)/V
        } else {
            beta[v] <- beta[v]
        }
    }
    phi[k, ] <- rdirichlet(1, beta)
}

dim(phi)

# check to make sure topic specific disease highly loaded
which(order(phi[1,],decreasing = T)%in%topic_spec_disease[1,])

## decide if we want all his or her diseases to come from same topic, no repeat diangoses
# disease <- list() # Initialize the list outside the loop
# for(i in 1:M){
#   disease[[i]] <- vector("list", Nd[i]) # Initialize the ith list
#   for(j in 1:Nd[i]){
#     zij <- which(rmultinom(1, size = 1, prob = theta[i,]) == 1)
#     wij <- which(rmultinom(1, 1, prob = phi[zij,]) == 1)
#     disease[[i]][j] <- wij # diagnose your patient with some badness 
#   }
# }


# recall we also want no diseases to come from same topic
disease <- list()  # Initialize the list outside the loop

for (i in 1:M) {
  # Check if the patient has any diagnoses to assign
  if (Nd[i] > 0) {
    disease[[i]] <- vector("list", Nd[i])  # Initialize the ith list if diagnoses are to be assigned
    assigned_diagnoses <- integer(0)  # Keep track of diagnoses already assigned to this patient

    for (j in 1:Nd[i]) {
      zij <- which(rmultinom(1, size = 1, prob = theta[i,]) == 1)
    
      # Exclude already assigned diagnoses when selecting a new one
      valid_diagnoses <- setdiff(1:V, assigned_diagnoses)
      
      # normalize so sums to 1
      prob_valid_diagnoses <- phi[zij, valid_diagnoses] / sum(phi[zij, valid_diagnoses])

      if (length(valid_diagnoses) > 0) {
        wij <- sample(valid_diagnoses, 1, prob = prob_valid_diagnoses)
        disease[[i]][[j]] <- wij
        assigned_diagnoses <- c(assigned_diagnoses, wij)  # Update the list of assigned diagnoses
      } else {
        # Handle the case where all diagnoses have been assigned
        # Placeholder: Assign a random diagnosis or handle as per your model's requirements
        disease[[i]][[j]] <- sample(1:V, 1)
      }
    }
  } else {
    disease[[i]] <- NULL  # Assign NULL or an appropriate placeholder for patients with no diagnoses
  }
}

```

# Visualizing Basic Simulation

Here we whow the distirbtuion of $theta$ and $phi$ to demonstrate that individuals are minimally loaded on one topic and each topic is minimally loaded on a few diseases:

```{r,echo=F}
ggplot(data = melt(theta),aes(x=value,fill=value,group=value))+geom_histogram()+labs(x="Theta_m",y="Density")+theme_classic()

ggplot(data = melt(phi),aes(x=value,fill=value,group=value))+geom_histogram()+labs(x="Phi_k",y="Density")+theme_classic()
```

But overall, `we can see that topic specific diseases have a higher $phi$ than those that don't:

```{r,echo=F}
par(mfrow=c(1,2))
hist(colSums(phi),main="Phi:all diseases",freq = F,nclass=20,col="red",xlim=c(0,0.5))
hist(colSums(phi[,topic_spec_disease]),main="Phi: topic spec diseases",freq = F,nclass=20,col="blue",xlim=c(0,0.5))

```

# Observed data

Now, we want to simulate the observed diagnoses $w_{ij}$ for sample individuals

```{r}
# Assuming 'disease' is your list of diagnoses for each person from above
# Convert 'disease' list into a data frame for plotting

disease_data <- data.frame(person = rep(seq_along(disease), sapply(disease, length)),
                        disease = unlist(disease))

# Now create the plot
random_peeps=sample(M,size = 5)
ggplot(disease_data[disease_data$person%in%random_peeps,], aes(x = disease,fill=as.factor(disease))) +
  geom_bar() +
  facet_wrap(~ person, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Disease", y = "Frequency", title = "Disease Distribution per Person",fill="Disease")+theme_classic()
```

We can see that most people have on average 10 diagnoses from a set of V = `r V`.
The people we choose, `r random_peeps`, were heavily loaded on the following topics:

`r sapply(random_peeps,function(x){which.max(theta[x,])})`

For example, our first person `r random_peeps[1]` has `r Nd[random_peeps[1]]` diagnoses. He is most heavily loaded on `r which.max(theta[random_peeps[1],])`. In turn this topic was chosen to have concentration of $phi_{k}$ at `r topic_spec_disease[which.max(theta[random_peeps[1],]),]` these disease. 

We can see that a high proportion of his diagnoses emanate from one of these topic-specific diseases:

`r length(intersect(unlist(disease[[random_peeps[1]]]),topic_spec_disease[which.max(theta[random_peeps[1],]),]))/d`.



# Adding a time element

Now, in McVean et al, instead of having only one V vector with N_{i} sparse indicators in $\phi_k$ per topic, there are $\phi_{kt}$ for a time and topic specific diagnosis distribution.


Easy enough! Recall before we had a matrix that was KxV for the disease distribution per topic.

Now we will initialize an *array* that is KxVxT that has a disease distribution per topic per time. 

The trick is that these need to be somewhat continuous -- you can't go from a 100% probability of hypertension to 0 overnight. 

So ... they use a natural spline. Here, we introduce a scaled bspline basis (with a knot at 0.5 and 3 degrees) that is scaled by the $\phi$ we just described inorder to introduce both continuity and preserve sparsity.


```{r}
K <- length(alpha)  # Number of topics
V <- length(beta)   # Number of diagnoses
T <- 100            # Number of time points
time_points <- seq(0, 100, length.out = T)
phi <- array(dim = c(K, V, T))

# Initialize phi with sparsity
for (k in 1:K) {
  for (v in 1:V) {
    # Assign higher probabilities to certain diseases for each topic
    if (v%in%topic_spec_disease[k,]) {
      phi[k, v, 1] <- 0.9  # Higher value for specific diseases
    } else {
      phi[k, v, 1] <- 0.1  # Lower value otherwise
    }
  }
}



# Normalize initial phi
for (k in 1:K) {
    phi[k, , 1] <- phi[k, , 1] / sum(phi[k, , 1])
}

summary(rowSums(phi[,,1]))

all_normalized <- all(apply(phi[, , 1], 1, function(x) abs(sum(x) - 1) < .Machine$double.eps^0.5))
print(all_normalized)  # Should print TRUE if all are normalized

degree=3
knots=0.5

# Apply spline transformation with positive coefficients
for (k in 1:K) {
    for (v in 1:V) {
        # Generate a B-spline basis
        bspline_basis <- bs(time_points, knots = knots, degree = degree, intercept = TRUE)
        
        # Generate positive coefficients for the spline
        coef <- runif(length(knots) + degree + 1, min = 0, max = 0.5)

        # Calculate spline values ensuring they are non-negative
        spline_values <- bspline_basis %*% coef

        # Scale the spline values based on initial phi value
        scaled_spline_values <- spline_values * phi[k, v, 1] / max(spline_values)

        # Store the scaled spline values
        phi[k, v, ] <- scaled_spline_values
    }
}

matplot(bspline_basis)
```

# Normalization

We need to normalize phi over time after this spline transfomration to ensure a proper distribution:
```{r}
# Normalize phi over time after spline transformation
for (t in 1:T) {
    for (k in 1:K) {
        sum_phi_kt <- sum(phi[k, , t])
        if (sum_phi_kt > 0) {
            phi[k, , t] <- phi[k, , t] / sum_phi_kt
        }
    }
}

# Check normalization for each time point
all_normalized <- TRUE
for (t in 1:T) {
    if (!all(apply(phi[, , t], 1, function(x) abs(sum(x) - 1) < .Machine$double.eps^0.5))) {
        all_normalized <- FALSE
        break
    }
}
print(all_normalized)  # Should print TRUE if all are normalized```
```

Now $\phi$ is the array of diseases x topics x time, so for a given topic, this should be continous. Let's look at a sample topic with a few diseases to check:

```{r, echo=F}
# Assuming phi is already created and is a KxVxT array
# Choose a topic to plot, for example, the first topic
selected_topic <- 2
selected_diseases=sample(V,size = 10)

# Extract the data for this topic
topic_data <- phi[selected_topic, ,]

# Create a sequence of time points for the x-axis
time_points <- seq(0, 10, length.out = dim(phi)[3])

# Plotting
matplot(time_points, t(topic_data), type = 'l', lty = 1,
        xlab = "Time", ylab = "Probability",
        main = paste("Disease Probabilities Over Time for Topic", selected_topic))
legend("topright", legend = paste("Disease", 1:dim(phi)[2]), col = 1:dim(phi)[2], lty = 1)
```


We can see that this corresponds with the disease we've chosen in topic_disease.

# Visualization part 2

In fact, if we want to visualize all diseases over time, we can consider the heatmap of $\phi_{k}$, and see that only the diseases that are part of the chosen topic specific disease have a non trivial coloring here.

```{r,echo=F}
library(ggplot2)
library(reshape2)

# Example: Creating a heatmap for Topic 1
# Extract the matrix for Topic 1 over time
phi_topic1 <- phi[1 ,, ]

# Convert to data frame
phi_topic1_df <- as.data.frame(t(phi_topic1))
names(phi_topic1_df) <- paste("Diagnosis", 1:V, sep = "_")
phi_topic1_df$Time <- 1:nrow(phi_topic1_df)

# Melt the data frame to long format
phi_topic1_long <- melt(phi_topic1_df, id.vars = "Time", variable.name = "Diagnosis", value.name = "Probability")

# Create heatmap
ggplot(phi_topic1_long, aes(x = Time, y= Diagnosis, fill = Probability)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(x = "Time", y = "Diagnosis", fill = "Probability") +
  ggtitle("Disease Distribution in Topic 1 Over Time") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## However this assumes that an individual stays within a topic

* Assumption: that the change in disease distribtuion for a given topic captures all the heterogeneity of disease coding
* We know that is not the case from our previous work: an individual can switch to a new pattern based on genetics and environmental factors, and we should have an ability to 'update' an individuals' topic distribtuion

* Furthermore, genetics has a *declining* importance over time, and the new diagnoses carry more weight. The rate of advancement *through a topic* (i.e., the shape of the spline) also may be affected by these intrinsic factors

* Finally, occupancy in one topic necessarily may prevent other topics

# Example function to update theta based on new diagnosis

Let's start by modeling an individual's transition to a new topic distribtuion as a funciton of genetics.

Here, we will have the genetics influence change by a decau factor, and the updated theta be proportional to the likelihood of the new diagnosis:

$\theta_{t+1}=p(D|\theta_{t})*p(\tilde{\theta})$

Here, $p(\tilde{\theta}$ represents the genetics-adapted prior distribtuion.


```{r}
update_theta <- function(current_theta, new_diagnosis, phi_time_dependent, genetics_influence, current_time, max_time) {
  # Ensure valid inputs
  if (any(is.na(current_theta)) || any(is.na(phi_time_dependent))) {
    return(current_theta)
  }

  # Decay factor for genetics influence
  decay_factor <- (max_time - current_time) / max_time
  adjusted_genetics_influence <- genetics_influence * decay_factor

  # Select phi for the current time
  phi_current_time <- phi_time_dependent[,,current_time]

  # Check for zero or NA in row sums
  row_sums <- rowSums(phi_current_time)
  if (any(row_sums == 0, is.na(row_sums))) {
    # Handle zero or NA row sums
    row_sums[row_sums == 0 | is.na(row_sums)] <- 1e-10
  }

  # Adjusted prior and likelihood
  adjusted_prior <- current_theta * adjusted_genetics_influence
  likelihood <- phi_current_time[, new_diagnosis] / row_sums

  # Bayesian update
  updated_theta <- likelihood * adjusted_prior
  total <- sum(updated_theta)
  
  if (total == 0 || is.na(total)) {
    return(current_theta)
  }
  return(updated_theta / total)  # Normalize the updated theta
}
```

Simulate data:

```{r}
M <- 100          # Number of patients
K <- 5            # Number of topics
V <- 300          # Number of diseases
T <- 100          # Number of time steps
max_time <- T     # Assuming T is the maximum time

# Initial topic distributions (theta) for each patient
initial_theta <- rdirichlet(M, shape = rep(0.1, K))

# Time-varying disease distributions within topics (phi)
# Assuming phi is a 3D array (topics x diseases x time)
# This needs to be defined based on your model

# Genetics influence (for simplicity, assuming a constant vector for each patient)
genetics_influence <- matrix(runif(M * K, 0.5, 1.5), nrow = M, ncol = K)

# Structure to hold theta values over time for each patient
theta_over_time <- array(dim = c(T, M, K))
theta_over_time[1, , ] <- initial_theta

## check to make sure all 1
all.equal(rowSums(theta_over_time[1,,]),rep(1,nrow(theta_over_time[1,,])))

# Function to simulate a new diagnosis (placeholder)
simulate_new_diagnosis <- function() {
  # Randomly simulate a new diagnosis
  return(sample(1:V, 1))
}


for (time in 2:T) {
  for (patient in 1:M) {
    # Simulate a new diagnosis for this patient at this time
    new_diagnosis <- simulate_new_diagnosis()

    # Update the patient's topic distribution based on the new diagnosis
    theta_over_time[time, patient, ] <- update_theta(
      current_theta = theta_over_time[time - 1, patient, ],
      new_diagnosis = new_diagnosis,
      phi_time_dependent = phi,  # Assuming phi is defined as before
      genetics_influence = genetics_influence[patient, ],
      current_time = time,
      max_time = max_time
    )
  }
}

## should  all be 1
rowSums(theta_over_time[2,,])

```

# Topic distribtuion per patient

Now let's select for a sample patient and view his topic distribution over time. 

```{r,echo=F}
# Select data for a specific patient, e.g., patient 1
pt=sample(M,size = 1)
theta_patient1 <- theta_over_time[,pt , ]

# Convert to a data frame and reshape
theta_patient1_df <- as.data.frame(theta_patient1)
colnames(theta_patient1_df) <- paste("Topic", 1:K, sep = "")
theta_patient1_df$Time <- 1:nrow(theta_patient1_df)

theta_patient1_long <- melt(theta_patient1_df, id.vars = "Time", variable.name = "Topic", value.name = "Proportion")


ggplot(theta_patient1_long, aes(x = Time, y = Proportion, fill = Topic)) +
  geom_area(position = 'stack') +
  labs(x = "Time", y = "Topic Proportion", fill = "Topic") +
  ggtitle("Distribution of Topics for Patient 1 Over Time")
```

## TODO: affect speed based through transitions based on genetics

```{r,eval=F}
# Pseudocode - Adjusting spline coefficients based on a regression model

# Assume a regression model is built to predict coefficients from genetic data
# coef_model <- lm(coef ~ genetic_factors, data = training_data)

# For each individual
for (i in 1:N_individuals) {
  # Get individual's genetic data
  genetic_data_i <- get_genetic_data_for_individual(i)

  # Predict/adjust coefficients based on genetic data
  predicted_coef <- predict(coef_model, newdata = genetic_data_i)

  # Use these coefficients for the spline
  bspline_basis <- bs(time_points, knots = knots, degree = degree, intercept = TRUE)
  spline_values <- bspline_basis %*% predicted_coef
  phi_individual[i, , ] <- spline_values
}
```

