---
title: "compute_at_risk_overtime"
output: html_document
date: "2023-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

## R Markdown

We show the risk for various starting points over time using the model which achieved minimal RMSE

```{r modelfit}
source("~/dynamichr/code/utils.R")
source("~/multistate2//code/smoothtest.R")
source("~/multistate2//code/newsmooth.R")
source("~/multistate2/code/fitarray.R")
library("reshape2")
source("~/multistate2/code/arrayindicate.R")

load("~/Library/CloudStorage/Dropbox-Personal///pheno_dir/output/merged_pheno_censor_final_withdrugs_smoke.rds")
dfh$cad.prs.lec=cut(dfh$cad.prs,breaks = c(-5,-0.84,0.84,5),labels = c("low","mid","high"))
dfh$int=interaction(dfh$f.31.0.0,dfh$cad.prs.lec)
# Relabel the levels of the interaction variable
levels(dfh$int) <- c(1,2,3,4,5,6)


train=dfh[1:(nrow(dfh)*0.80),]

dfascvd=readRDS("~/multistate2//output/dfascvd_newbp.rds")
test=dfh[!(dfh$identifier%in%train$identifier),]
test=merge(test,dfascvd[,-which(names(dfascvd)%in%c("age","anylipidmed0","bp_med2","smoke"))],by.x="identifier",by.y="sample_id")
test$ascvd_10y_accaha=test$as2
test$phenos.enrollment=test$f.21003.0.0
test=data.table(test)
source("~/multistate2/code/frs30_URBUT/fun.frs_30y.R")
library(CVrisk)
ages=c(40:80)
nstates=c("Health", "Ht","HyperLip","Dm","Cad","death","Ht&HyperLip","HyperLip&Dm","Ht&Dm","Ht&HyperLip&Dm")

 modelfit = fitfunc2(
   data.table(train),
   ages = ages,
    nstates = nstates,
 mode = "binomial",
   covariates = "cad.prs+f.31.0.0+smoke+antihtn_now+statin_now"
 )

# B=50
# ## increase bootstrap
# 
# for(i in 1:B){
#   
# row=sample(nrow(train),size = nrow(train),replace = T)
# modelfit = fitfunc2(
#   data.table(train[row,]),
#   ages = ages,
#   nstates = nstates,
#   mode = "binomial",
#   covariates = "cad.prs+f.31.0.0+smoke+antihtn_now+statin_now")
# 
# saveRDS(object = modelfit,file = paste0("~/multistate2/output/bootstrap",i,".rds"))
# }
ages = 40:80
```


## First we plot for some sample individuals

```{r}
## return a matrix of coefficients over all ages for a given state to state transition
a = coefplotsmooth2(
  ages = ages,
  start = "Health",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
healthy_coefs = a$custom_smooth

b = coefplotsmooth2(
  ages = ages,
  start = "Ht",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 30,
  span = 0.75,
  degree = 2
)
ht_coefs = b$custom_smooth

c = coefplotsmooth2(
  ages = ages,
  start = "HyperLip",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
hyperlip_coefs = c$custom_smooth

d = coefplotsmooth2(
  ages = ages,
  start = "Dm",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
dm_coefs = d$custom_smooth


e = coefplotsmooth2(
  ages = ages,
  start = "Ht&HyperLip",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
HH_coefs = e$custom_smooth

f = coefplotsmooth2(
  ages = ages,
  start = "HyperLip&Dm",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
HD_coefs = f$custom_smooth

g = coefplotsmooth2(
  ages = ages,
  start = "Ht&Dm",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)
TD_coefs = g$custom_smooth

h = coefplotsmooth2(
  ages = ages,
  start = "Ht&HyperLip&Dm",
  stop = "Cad",
  modelfit = modelfit,
  window_width = 20,
  span = 0.75,
  degree = 2
)

HHD_coefs = h$custom_smooth

coeflist = list(
  healthy_coefs,
  ht_coefs,
  hyperlip_coefs,
  dm_coefs,
  HH_coefs,
  HD_coefs,
  TD_coefs,
  HHD_coefs
)

agesint = seq(40, 79, by = 1)
predicted_risks = array(NA, dim = c(nrow(test), length(agesint), 8))
dimnames(predicted_risks) = list(
  as.numeric(as.character(test$identifier)),
  agesint,
  c(
    "Health",
    "Ht",
    "HyperLip",
    "Dm",
    "Ht&HyperLip",
    "HyperLip&Dm",
    "Ht&Dm",
    "Ht&HyperLip&Dm"
  )
)

for (j in 1:length(coeflist)) {
  coefs = coeflist[[j]]
  
  for (i in 1:length(agesint)) {
    age = agesint[i]
    start = age
    stop = 80
    atrisk = test
    ar = data.frame(
      "intercept" = 1,
      "cad.prs" = atrisk$cad.prs,
      "sex" = atrisk$f.31.0.0,
      "smoke" = atrisk$smoke,
      "antihtn_now" = ifelse(atrisk$antihtn == 1 &
                               atrisk$htn_age < age, 1, 0),
      "statin_now" = ifelse(atrisk$statin == 1 &
                              atrisk$statin_age < age, 1, 0)
      
    )
    
    
    mso = compute_prediction_product_matrix(
      coefmat = coefs,
      atrisk = ar,
      agepredinterval = c(start:stop)
    )
    
    predicted_risks[, i, j] = mso$PredictedIntervalrisk
  }
  
}



ipf = function(i) {
  m = melt(predicted_risks[i, , ])
  atrisk = test
  ar = data.frame(
    "intercept" = 1,
    "cad.prs" = atrisk$cad.prs,
    "sex" = atrisk$f.31.0.0,
    "smoke" = atrisk$smoke,
    "antihtn_now" = ifelse(atrisk$antihtn == 1,
                           atrisk$htn_age, 0),
    "statin_now" = ifelse(atrisk$statin == 1,
                          atrisk$statin_age, 0)
    
  )
  g = ggplot(m, aes(Var1, value, col = Var2)) + geom_point() +
    labs(x = "Age of consideration", y = "Lifetime Risk", col = "Starting State") +
    theme_classic() + ggtitle(paste0(
      "PRS:",
      round(ar[i, 2], 0),
      ",Sex:",
      ar[i, 3],
      ",Smoke:",
      ar[i, 4],
      ",AH:",
      round(ar[i, 5], 0),
      "Stat:",
      round(ar[i, 6], 0)
    ))
  
  ggplotly(g)
}

ipf(1)
ipf(18)
ipf(23)
ipf(30)

```

For Fixed at risk matrix

```{R}
intercept=1
cad.prs=c(-2,-1,0,1,2)
sex=c(0,1)
smoke=c(0,1)
antihtn_now=c(0,1)
statin_now=c(0,1)

atrisk=expand.grid(intercept,cad.prs,sex,smoke,antihtn_now,statin_now)



agesint=seq(40,70,by=1)
predicted_risks2=array(NA,dim=c(nrow(atrisk),length(agesint),8))
predicted_risks_treat=array(NA,dim=c(nrow(atrisk),length(agesint),8))
dimnames(predicted_risks2)=list(rownames(atrisk),agesint,c("Health","Ht","HyperLip","Dm","Ht&HyperLip","HyperLip&DM","Ht&Dm","Ht&HyperLip&Dm"))

for(j in 1:length(coeflist)){
  coefs=coeflist[[j]]
  
  for(i in 1:length(agesint)){
    age=agesint[i]
    start=age
    stop=80
    ar=atrisk
    

  mso = compute_prediction_product_matrix(
    coefmat = coefs,
    atrisk = ar,
    agepredinterval = c(start:stop)
  )
  
  predicted_risks2[,i,j] = mso$PredictedIntervalrisk
  predicted_risks_treat[,i,j] = mso$Hazard_treated
  }
  
}

## also data_array on the app
saveRDS(predicted_risks2,"~/output/predictedrisk_forfixedmatrix.rds")

ipf2=function(i){
m=melt(predicted_risks2[i,,])
ar=atrisk
g=ggplot(m,aes(Var1,value,col=Var2))+geom_point()+
  labs(x="Age of consideration",y="Lifetime Risk",col="Starting State")+
  theme_classic()+ggtitle(paste0("PRS:",round(ar[i,2],0),",Sex:",ar[i,3],",Smoke:",ar[i,4],",AH:",round(ar[i,5],0),"Stat:",round(ar[i,6],0)))
 
ggplotly(g)}

## fix, new PRS
ipf(i=1)
ipf2(i=5)

## fix pRS, new sex
ipf(i=1)
ipf2(i=6)

## fix pRS, new smoke
ipf(i=1)
ipf2(i=11)

## fix pRS, new AH
ipf(i=1)
ipf2(i=21)


## fix pRS, new stat
ipf(i=1)
ipf2(i=41)
```

Investigate overlap:


```{r}
#install.packages("VennDiagram")   # Install & load VennDiagram package
library("VennDiagram")

ht=fread("~/Desktop/output_gp/Hypertension.tab.tsv.gz")
hyperlip=fread("~/Desktop/output_gp/Hypercholesterolemia.tab.tsv.gz")
cad=fread("~/Desktop/output_gp/Coronary_Artery_Disease_INTERMEDIATE.tab.tsv.gz")
dm=fread("~/Desktop/output_gp/Diabetes_Type_2.tab.tsv.gz")
g=merge(dfh[c("identifier","Ht_0_Any","HyperLip_0_Any","Cad_0_Any","antihtn","statin")],ht[,c("sample_id","has_disease")],by.x="identifier",by.y="sample_id")
g2=merge(g,hyperlip[,c("sample_id","has_disease")],by.x="identifier",by.y="sample_id")
g3=merge(g2,cad[,c("sample_id","has_disease")],by.x="identifier",by.y="sample_id")
g4=merge(dfh[,c("identifier","Dm_0_Any")],dm[,c("sample_id","has_disease")],by.x="identifier",by.y="sample_id")


pdf("../output/hypertense.pdf")
plot(euler(list("UkbPheno"=which(g2$Ht_0_Any==2),"Lab"=which(g2$has_disease.x==1),"AntiHtn"=which(g2$antihtn==1))),main="Hypertenson",quantities = T)
dev.off()


pdf("../output/hyperlip.pdf")
plot(euler(list("UkbPheno"=which(g2$HyperLip_0_Any==2),"Lab"=which(g2$has_disease.y==1),"Statin"=which(g2$statin==1))),main="HyperLipidemia",quantities = T)
dev.off()


pdf("../output/cad.pdf")
plot(euler(list("UkbPheno"=which(g3$Cad_0_Any==2),"Lab"=which(g3$has_disease==1),"Statin"=which(g3$statin==1))),main="CAD",quantities = T)
dev.off()


pdf("../output/dm.pdf")
plot(euler(list("UkbPheno"=which(g4$Dm_0_Any==2),"Lab"=which(g4$has_disease==1))),main="DM",quantities = T)
dev.off()
```


```{r,eval=T}

get_risk_profile <- function(data_array, 
                             start_cov_profile, change_age_cov = NULL, new_cov_profile = NULL, 
                             start_model, change_ages_model = NULL, new_models = NULL) {
  
  # Validate if the provided input values are within the range
  if(start_cov_profile < 1 || start_cov_profile > 50) stop("Invalid starting covariate profile.")
  if(!is.null(new_cov_profile) && (new_cov_profile < 1 || new_cov_profile > 50)) stop("Invalid new covariate profile.")
  if(start_model < 1 || start_model > 8) stop("Invalid starting model.")
  
  # Placeholder to store the extracted risk values
  risk_profile <- numeric(length(40:70))
  
  age_range <- 40:70
  for(i in 1:length(age_range)) {
    age <- age_range[i]
    current_cov <- start_cov_profile
    current_model <- start_model
    
    # Check if we need to change the covariate profile based on the age
    if(!is.null(change_age_cov) && age >= change_age_cov) {
      current_cov <- new_cov_profile
    }
    
    # Check if we need to change the model based on the age
    if(!is.null(change_ages_model)) {
      # Assuming change_ages_model and new_models are vectors of the same length
      change_idxs <- which(age >= change_ages_model)
      if(length(change_idxs) > 0) {
        current_model <- new_models[max(change_idxs)]
      }
    }
    
    # Extract risk value from the data array
    risk_profile[i] <- data_array[current_cov, i, current_model]
    #print(i)
  }
  
  return(risk_profile)
}

# Example usage
# Let's assume your 3D array is named data_array
# risk_values <- get_risk_profile(data_array, start_cov_profile = 5,
# change_age_cov = 55, new_cov_profile = 10,
# start_model = 2, change_ages_model = c(50, 60), new_models = c(5, 6))

## this is not for boostratpped
saveRDS(predicted_risk,"../output/predicted_risks_array_fulltest.rds")
s=statusarray(df_frame = data.table(test),ages = ages,nstates = nstates)
s2=s[c(1:40),,c(1,2,3,4,7,8,9,10)]

```

```{r}
mean(s2["40",,"Health"])
mean(s2["40",,"Ht"])
mean(s["40",,"Cad"])


mean(s2["76",,"Health"])
mean(s2["40",,"Ht"])
mean(s["40",,"Cad"])

```{r,eval=F}
# Get dimensions
num_individuals <- length(rownames(predicted_risks))
num_ages <- length(ages) - 1

# Initialize the result matrix with zeroes
cr <- matrix(0, nrow=num_individuals, ncol=num_ages)

# Iterate over individuals and ages
for(i in 1:num_individuals){
  individual <- as.character(rownames(predicted_risks)[i])
  
  for(j in 1:num_ages){
    agename <- as.character(ages[j])
    
    status <- which(s2[agename, individual,] == 1)
    
    if(length(status) > 0) {
      cr[i,j] <- as.numeric(as.character(predicted_risks[individual, agename, status]))
    } else {
      # This will help you debug if there's an issue with particular combinations
      cr[i,j] =0
      #cat(sprintf("No status found for individual %s at age %s\n", individual, agename))
    }
  }
}


saveRDS(cr,"~/multistate2/output/cr3.rds")
```

```{r,echo=F}
plotfunc=function(patientriskmat,patient,ages){
  riskvec=patientriskmat[patient,]
  
  mat=data.frame(ages=ages,lifetimerisk=100*riskvec)
  
  mat=mat[mat$lifetimerisk>0,]
  mat$risklev=cut(mat$lifetimerisk,breaks = c(0,10,15,20,25,30,35,40,50,100))
  ggplot(mat[mat$lifetimerisk>0,],aes(ages,lifetimerisk,col=risklev))+geom_point()+labs(x="Ages",y="Lifetime Risk",col="Lifetime Risk Level")
}

```


Using our multistate approach MSGene, we describe the overall state distribution across the lifespan in our cohort, normalizing to exclude censoring at each age (Figure 2). At age 40, 94.4% of individuals are in the healthy category, with 4.1% in the hypertensive category and 0.3% with a diagnosis of CAD. By age 76, CAD occupancy peaks at 12.5% of uncensored individuals and health declines to 27.6% of uncensored individuals. Death dominates as a competing risk with time, with 7.4% of individuals having died by age 80. 

# look at h

```{r}
cr=readRDS("~/multistate2/output/cr3.rds")
h=apply(s[,,c(1:10)],3,function(x){colSums(t(x))})
h=apply(h,1,function(x){x/sum(x)})
h=t(h)
melt=melt(h)
g=ggplot(melt[!(melt$Var2 %in% "out"),],aes(Var1,value,fill=Var2))+geom_bar(stat="identity")+
  labs(x="Age",y="Proportional Occupancy",fill="Current State")+scale_fill_d3()

saveRDS(g,file="~/multistate2/output/proportionaloccupancy.rds")

plotfunc(patientriskmat = cr,patient = which(test$identifier%in%"5021941"),ages = c(40:79))
```


```{r}
test$sta_lev=cut(as.numeric(test$statin_age,breaks = c(0,40,50,60,70,100)))



get_person_summary_plot <- function(pm, s2, person_idx) {
  # Convert the risk array for the given person to a data frame
  risk_df <- as.data.frame(as.table(pm[, person_idx, , ]))
  colnames(risk_df) <- c("Bootstrap", "Age", "State", "Risk")

  # Convert the state indicator array for the given person to a data frame
  state_df <- as.data.frame(as.table(s2[, person_idx, ]))
  colnames(state_df) <- c("Age", "State", "Indicator")

  # Extract the state info for the given person
  current_state_df <- state_df %>%
    filter(Indicator == 1) %>%
    dplyr::select(-Indicator)
  
  # Merge the data to get risks based on the state the person is in
  merged_df <- risk_df %>%
    inner_join(current_state_df, by = c("Age", "State"))
  
  # Compute mean and SE for the person across ages
  summary_df <- merged_df %>%
    group_by(Age) %>%
    summarise(
      mean_risk = mean(Risk),
      se_risk = sd(Risk) / sqrt(n()),
      lower_95 = mean_risk - 1.96 * se_risk,  # For 95% CI
      upper_95 = mean_risk + 1.96 * se_risk,  # For 95% CI
      #lower_95 = quantile(Risk,0.025),  # For 95% CI
      #upper_95 = quantile(Risk,0.975),  # For 95% CI
    
      .groups = "drop"
    )

  # Add a risk level column to the data frame
  summary_df$risk_level <- cut(summary_df$mean_risk*100, breaks = c(-Inf, 20, 30, Inf), labels = c("<20", "20-30", ">30"))
  summary_df$risk_level=as.factor(summary_df$risk_level)
  # Create ggplot
  plot <- ggplot(summary_df, aes(x = Age, y = mean_risk*100,color=as.factor(risk_level))) +
    geom_point(aes(color = as.factor(risk_level)), size = 3) +scale_color_d3()+
    geom_errorbar(aes(ymin = lower_95*100, ymax = upper_95*100), width = 0.3) +
    
    labs(title = paste("Predicted Risk by Ages for Person", person_idx),
         y = "Remaining Lifetime Risk (%)", x = "Age",col="Risk Level") +
    theme_classic(base_size = 20)

  return(list(plot,summary_df))
}

# Call the function for the first person and display the plot:
get_person_summary_plot(pm, s2, "5021941")



get_person_cross_threshold <- function(pm, s2, person_idx) {
  # Convert the risk array for the given person to a data frame
  risk_df <- as.data.frame(as.table(pm[, person_idx, , ]))
  colnames(risk_df) <- c("Bootstrap", "Age", "State", "Risk")

  # Convert the state indicator array for the given person to a data frame
  state_df <- as.data.frame(as.table(s2[, person_idx, ]))
  colnames(state_df) <- c("Age", "State", "Indicator")

  # Extract the state info for the given person
  current_state_df <- state_df %>%
    filter(Indicator == 1) %>%
    select(-Indicator)
  
  # Merge the data to get risks based on the state the person is in
  merged_df <- risk_df %>%
    inner_join(current_state_df, by = c("Age", "State"))
  
  # Compute mean and SE for the person across ages
  summary_df <- merged_df %>%
    group_by(Age) %>%
    summarise(
      mean_risk = mean(Risk),
      se_risk = sd(Risk) / sqrt(n()),
      #lower_95 = mean_risk - 1.96 * se_risk,  # For 95% CI
      #upper_95 = mean_risk + 1.96 * se_risk,  # For 95% CI
      lower_95 = quantile(Risk,0.025),  # For 95% CI
      upper_95 = quantile(Risk,0.975),  # For 95% CI
      num_exceeded=sum(Risk>thresh),
    
      .groups = "drop"
    )

return(list(mean_risk=summary_df$mean_risk,se=summary_df$se_risk,num_exceeded=summary_df$num_exceeded))}

states=sapply(c(1:nrow(test)),function(x){get_person_cross_threshold(pm,s2,as.character(test$identifier[x]))$mean_risk})

saveRDS(states,"~/multistate2/output/state_occupancy_risk.rds")

##


thresh=0.10;
t=sapply(1:length(states),function(x){w=which(states[[x]]>thresh);ifelse(length(w)>0,min(ages[w]),0)})

library(ggplot2)

# Assuming t is your data vector...
data <- data.frame(t_val = t[t > 0])

# Create the ggplot histogram
p <- ggplot(data, aes(x = t_val)) +
  geom_histogram(bins = length(ages) - 1, fill = "blue", color = "black", alpha = 0.7, aes(y =after_stat(density))) +
  labs(title = "Histogram of First Age Exceeding Risk",
       x = "Minimum Age Exceeded",
       y = "Density")+theme_classic()

# Display the plot
print(p)

# Sample data (replace this with your matrix or data frame)

data=test
data$statin_age=as.numeric(data$statin_age)
# Categorize the ages
data$age_category <- case_when(
  is.na(data$statin_age) ~ "Never",
  data$statin_age < 45 ~ "<45",
  data$statin_age >= 45 & data$statin_age < 55 ~ "45-55",
  data$statin_age >= 55 & data$statin_age < 65 ~ "55-65",
  data$statin_age >= 65 ~ ">65",
  TRUE ~ as.character(data$statin_age) # catch-all for unexpected values
)

# Order the levels of the age_category factor
data$age_category <- factor(data$age_category, levels = c("Never", "<45", "45-55", "55-65", ">65"))


# Prepare data for plotting
plot_data <- data %>%
  group_by(age_category) %>%
  arrange(cad.prs) %>%
  mutate(ecdf_val = ecdf(cad.prs)(cad.prs))


# Prepare data for plotting
plot_data2 <- data %>%
  group_by(age_category) %>%
  arrange(cad.prs)


statage = ggplot(plot_data, aes(x = cad.prs, y = ecdf_val, color = age_category)) +
  geom_step(size = 1.5) +  # Setting the line size for the plot
  labs(title = "eCDF of PRS by Statin Start Age",
       x = "PRS",
       y = "eCDF",
       color = "First Age (prescription or diagnosis)") + scale_color_aaas() +
  theme_classic(base_size = 20) + theme(legend.position = "none")


###

# Sample data (replace this with your matrix or data frame)

data=test
data$Cad_0_censor_age=as.numeric(data$Cad_0_censor_age)
# Categorize the ages
sickdata=data[data$Cad_0_Any==2,]
sickdata$age_category<- case_when(
  #is.na(data$statin_age) ~ "Never",
  sickdata$Cad_0_censor_age < 45 ~ "<45",
  sickdata$Cad_0_censor_age >= 45 & sickdata$Cad_0_censor_age  < 55 ~ "45-55",
  sickdata$Cad_0_censor_age >= 55 & sickdata$Cad_0_censor_age  < 65 ~ "55-65",
  sickdata$Cad_0_censor_age >= 65 ~ ">65",
  TRUE ~ as.character(sickdata$Cad_0_censor_age) # catch-all for unexpected values
)

# Order the levels of the age_category factor
sickdata$age_category <- factor(sickdata$age_category, levels = c("<45", "45-55", "55-65", ">65"))

health_data=test[Cad_0_Any==1,]
health_data$age_category="Never"

d2=rbind(health_data,sickdata)
# Prepare data for plotting
plot_data <- d2 %>%
  group_by(age_category) %>%
  arrange(cad.prs) %>%
  mutate(ecdf_val = ecdf(cad.prs)(cad.prs))



cadage = ggplot(plot_data, aes(x = cad.prs, y = ecdf_val, color = age_category)) + geom_step(size = 1.5) +  # Setting the line size for the plot
  labs(title = "eCDF of PRS by CAD Diag Age",
       x = "PRS",
       y = "eCDF",
       color = "First Age (prescription or diagnosis)") + scale_color_aaas() + theme_classic(base_size = 20) + theme(legend.position = "none") +guides(color = guide_legend(override.aes = list(size = 1.5)))  # Setting the line size for the legend


g=ggarrange(statage,cadage)
ggsave(g,file="../output/ecdf.pdf")
library(dplyr)

pm=readRDS("../output/predictedrsiskboot.rds")



get_person_exceedance_count <- function(person_index, pm, s2, threshold) {
  person_risks <- pm[, person_index, , ]
  person_states <- s2[, person_index, ]

  age_count <- dim(pm)[3]
  
  # Placeholder for results
  result <- data.frame(Age = as.numeric(dimnames(pm)[3][[1]]), Count_Exceedances = integer(age_count))
  
  for(age in 1:age_count) {
    current_state <- which(person_states[age, ] == 1)
    exceedances <- sum(person_risks[, age, current_state] > threshold)
    result$Count_Exceedances[age] <- exceedances
  }
  
  return(result)
}

# Test for a single person
threshold <- 0.1  # set to your desired threshold
person_index = 1  # example person
result <- get_person_exceedance_count(person_index, pm, s2, threshold)
print(result)

threshold <- 0.1 

temp=proc.time()
# Set to your desired threshold
all_results=lapply(1:dim(pm)[2], function(i) {
  df <- get_person_exceedance_count(i, pm, s2, threshold)
  df$Individual <- i
  return(df)
}))

proc.time()-temp

s=t(sapply(1:length(l),function(x){l[[x]]$Count_Exceedances}))

result_df <- as.data.frame(s)
result_df$Individual <- 1:nrow(s)
long_result_df <- result_df %>%
  gather(Age, Count_Exceedances, -Individual)


long_result_df$NumericAge <- as.numeric(gsub("V", "", long_result_df$Age))

# Compute mean exceedances per age
age_summary <- long_result_df %>%
  group_by(NumericAge) %>%
  summarise(MeanExceedances = sum(Count_Exceedances), .groups = 'drop')

# Ensure all ages are represented
all_ages <- data.frame(NumericAge = 40:79)
age_summary <- merge(all_ages, age_summary, by = "NumericAge", all.x = TRUE)
age_summary$MeanExceedances[is.na(age_summary$MeanExceedances)] <- 0

# Plot
ggplot(age_summary, aes(x = NumericAge, y = MeanExceedances)) +
  geom_col(fill = "dodgerblue", color = "black") +
  labs(title = "Mean Number of Simulations Exceeding Threshold by Age",
       x = "Age", 
       y = "Mean Number of Simulations Exceeding Threshold") +
  theme_minimal()




threshold=0.10
# Create an array to store the exceedance data (1 if exceeded, 0 otherwise)

# Create an array to store the exceedance data (1 if exceeded, 0 otherwise)
exceedance_array <- array(0, dim = c(dim(s2)[1:2]))

threshold=0.20
# Check for exceedance in each bootstrap, for each patient's state at each age
for(b in 1:dim(pm)[1]){
  for(i in 1:dim(pm)[2]){
    for(a in 1:dim(pm)[3]){
      state_at_age <- which(s2[a, i, ] == 1)
      if (length(state_at_age) > 0) {
        if(pm[b, i, a, state_at_age] > threshold){
          exceedance_array[a, i] <- 1
        }
      }
    }
  }
}

dimnames(exceedance_array)[1][[1]]=c(40:79)

agenames=c(40:79)
# Extract the minimum age of exceedance for each patient
min_age_exceedance <- apply(exceedance_array, 2, function(x) {
  if(sum(x) > 0){
    return(agenames[which(x == 1)[1]])
  } else {
    return(NA)
  }
})

# Convert the result into a dataframe for plotting
df <- data.frame(MinAgeExceedance = min_age_exceedance)



ggplot(df, aes(x = MinAgeExceedance)) +
  geom_histogram(binwidth = 1, fill = "dodgerblue", color = "black") +
  labs(title = "Distribution of Minimum Age of Risk Exceedance",
       x = "Minimum Age of Exceedance (with boostrapping)", 
       y = "Number of Patients") +
  theme_minimal()


```