---
title: "ThresholdVariant"
format: html
editor: visual
---

# **Simulate a Continuous Genetic Risk Score**:

As before, we simulate the genetic risk score with a combination of rare and common variants.

```{r,echo=FALSE}
#install.packages("survival")
library(survival)

```

-   The distributions from which the effect sizes for rare and common variants are drawn should be chosen based on real-world data or literature to reflect actual genetic risk profiles.

-   The number of rare and common variants (**`n_rare_variants`** and **`n_common_variants`**) and their frequencies should be set to realistic values.

-   This approach results in a more nuanced genetic risk score, which is more representative of actual genetic variability in a population.

```{r}
n <- 1000  # Number of individuals
n_rare_variants <- 10  # Number of rare variants per individual
n_common_variants <- 50  # Number of common variants per individual

# Simulating the effect of rare variants
rare_effect_size <- rnorm(n_rare_variants, mean = 1.5, sd = 0.5)  # Adjust mean and sd as needed
rare_variants <- matrix(rbinom(n * n_rare_variants, 1, 0.01), ncol = n_rare_variants) * rare_effect_size

# Simulating the effect of common variants
common_effect_size <- rnorm(n_common_variants, mean = 0.1, sd = 0.05)  # Adjust mean and sd as needed
common_variants <- matrix(rbinom(n * n_common_variants, 1, 0.5), ncol = n_common_variants) * common_effect_size

# Combined genetic risk score
genetic_risk_score <- rowSums(rare_variants) + rowSums(common_variants)
```

**Simulate Event Times Using the Weibull Distribution**:

-   The choice of shape parameter (greater than 1) is crucial to ensure an increasing hazard rate over time, leading to a 'J'-shaped distribution.

-   The scale parameter, influenced by the genetic risk score, will affect the distribution of event times but won't change the overall shape of the hazard function.

-   The exact values for the shape and scale parameters, as well as the distributions for the genetic risk scores, should be based on the specific characteristics of the disease or trait you're studying.

```{r}

# Parameters for the Weibull distribution
shape <- 3  # Shape parameter > 1 for increasing hazard
scale_base <- 50  # Base scale parameter

# Adjust scale parameter based on genetic risk score
scale <- scale_base * exp(-0.03 * genetic_risk_score)

# Simulate event times
event_times <- rweibull(n, shape, scale)

# Adjust event times to a realistic age range (e.g., 20 to 80 years)
event_times <- 20 + (event_times / max(event_times)) * 60
```

**Analysis and Visualization**:

Analyze or visualize the distribution of event times to understand the impact of the genetic risk.

```{r}
hist(event_times, breaks = 30, main = "Distribution of Simulated Event Ages", xlab = "Age", ylab = "Frequency")
```
